### 2020-10-14 police_eda.Rmd
First commit

---
title: "Police EDA"
author: Connie Wu
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
```

## Load Data
```{r load-data}
all <- read.csv("allshootingsdata.csv")

philly <- read.csv("philly_shootings.csv")
cincinnati <- read.csv("cincinnati_shootings.csv")
```

## Exploratory Data Analysis

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

all <- all %>% 
  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")

officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
out <- which((officer_gender == officer_race) == FALSE)
all <- all[-out,]

all <- all %>% 
  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
  separate_rows(OfficerRace, OfficerGender, sep = ",")

officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
out2 <- which((officer_gender2 == officer_race2) == FALSE)
all <- all[-out2,]

all$OfficerRace <- trimws(all$OfficerRace)
all$OfficerGender <- trimws(all$OfficerGender)
all$Fatal <- trimws(all$Fatal)

all <- all %>% 
  filter(SubjectRace != "U") %>% 
  filter(OfficerRace != "U") %>% 
  filter(OfficerRace != "Unknown") %>% 
  filter(OfficerRace != "NA") %>% 
  filter(OfficerRace != "") %>% 
  filter(OfficerRace != "I") %>% 
  filter(OfficerGender != "U") %>% 
  filter(OfficerGender != "N") %>% 
  filter(SubjectGender != "U") %>% 
  filter(SubjectGender != "N") %>% 
  filter(Fatal != "U")

all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"

all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
all$OfficerGender[all$OfficerGender == "Male"] <- "M"
```

```{r fatal-eda}
temp <- all %>% 
  group_by(Fatal) %>% 
  summarise(count = n())

ggplot(temp, aes(x=Fatal, y=count)) + 
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
```

```{r race-eda}
temp <- all %>% 
  group_by(Fatal, SubjectRace) %>% 
  summarise(race_count = n())

ggplot(temp, aes(x=SubjectRace, y=race_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in Police Shootings")

temp2 <- all %>% 
  group_by(Fatal, OfficerRace) %>% 
  summarise(race_count = n())

ggplot(temp2, aes(x=OfficerRace, y=race_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
```

```{r gender-eda}
temp <- all %>% 
  group_by(Fatal, SubjectGender) %>% 
  summarise(gender_count = n())

ggplot(temp, aes(x=SubjectGender, y=gender_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in Police Shootings")

temp2 <- all %>% 
  group_by(Fatal, OfficerGender) %>% 
  summarise(gender_count = n())

ggplot(temp2, aes(x=OfficerGender, y=gender_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
```

```{r weapon-eda}
temp <- all %>% 
  group_by(Fatal, weapon) %>% 
  summarise(count = n())

ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings") 
```

```{r clean-cities}
cinn <- cincinnati %>% 
  select(RACE, SEX, TYPE) %>% 
  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
  mutate(city = "Cincinnati")

phil <- philly %>% 
  select(race, sex, fatal) %>% 
  rename(gender = sex) %>% 
  mutate(city = "Philadelphia")

comp <- rbind(cinn, phil)

comp <- comp %>%
  filter(race != "UNKNOWN") %>% 
  filter (race != "") %>% 
  filter(race != "UNK") %>% 
  filter(gender != "")

comp$race[comp$race == "WHITE"] <- "W"
comp$race[comp$race == "BLACK"] <- "B"
comp$race[comp$race == "OTHER"] <- "O"
comp$race[comp$race == "HISPANIC"] <- "H"

comp$gender[comp$gender == "MALE"] <- "M"
comp$gender[comp$gender == "FEMALE"] <- "F"

comp$fatal[comp$fatal == "0"] <- "N"
comp$fatal[comp$fatal == "1"] <- "F"
comp$fatal[comp$fatal == "NONFATAL"] <- "N"
comp$fatal[comp$fatal == "FATAL"] <- "F"
```

```{r fatal-eda}
temp <- comp %>% 
  group_by(fatal) %>% 
  summarise(count = n())

ggplot(temp, aes(x=fatal, y=count)) + 
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
```
```{r race-eda-cities}
temp <- comp %>% 
  group_by(fatal, race) %>% 
  summarise(count = n())

ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
```

```{r gender-eda-cities}
temp <- comp %>% 
  group_by(fatal, gender) %>% 
  summarise(count = n())

ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
```

### 2020-10-14 police_modeling.Rmd
First commit

--- 2020-10-14 police_eda.Rmd
@@
 ---
-title: "Police EDA"
-author: Connie Wu
+title: "Modeling"
+author: "Connie Wu"
+date: "10/13/2020"
 output: html_document
@@
 
+## Modeling
 ```{r}
@@
 
-## Load Data
 ```{r load-data}
@@
 ```
-
-## Exploratory Data Analysis
 
@@
 ```
-
-```{r fatal-eda}
-temp <- all %>% 
-  group_by(Fatal) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=Fatal, y=count)) + 
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-```
-
-```{r race-eda}
-temp <- all %>% 
-  group_by(Fatal, SubjectRace) %>% 
-  summarise(race_count = n())
-
-ggplot(temp, aes(x=SubjectRace, y=race_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in Police Shootings")
-
-temp2 <- all %>% 
-  group_by(Fatal, OfficerRace) %>% 
-  summarise(race_count = n())
-
-ggplot(temp2, aes(x=OfficerRace, y=race_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-```{r gender-eda}
-temp <- all %>% 
-  group_by(Fatal, SubjectGender) %>% 
-  summarise(gender_count = n())
-
-ggplot(temp, aes(x=SubjectGender, y=gender_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in Police Shootings")
-
-temp2 <- all %>% 
-  group_by(Fatal, OfficerGender) %>% 
-  summarise(gender_count = n())
-
-ggplot(temp2, aes(x=OfficerGender, y=gender_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-```{r weapon-eda}
-temp <- all %>% 
-  group_by(Fatal, weapon) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-```{r clean-cities}
-cinn <- cincinnati %>% 
-  select(RACE, SEX, TYPE) %>% 
-  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
-  mutate(city = "Cincinnati")
-
-phil <- philly %>% 
-  select(race, sex, fatal) %>% 
-  rename(gender = sex) %>% 
-  mutate(city = "Philadelphia")
-
-comp <- rbind(cinn, phil)
-
-comp <- comp %>%
-  filter(race != "UNKNOWN") %>% 
-  filter (race != "") %>% 
-  filter(race != "UNK") %>% 
-  filter(gender != "")
-
-comp$race[comp$race == "WHITE"] <- "W"
-comp$race[comp$race == "BLACK"] <- "B"
-comp$race[comp$race == "OTHER"] <- "O"
-comp$race[comp$race == "HISPANIC"] <- "H"
-
-comp$gender[comp$gender == "MALE"] <- "M"
-comp$gender[comp$gender == "FEMALE"] <- "F"
-
-comp$fatal[comp$fatal == "0"] <- "N"
-comp$fatal[comp$fatal == "1"] <- "F"
-comp$fatal[comp$fatal == "NONFATAL"] <- "N"
-comp$fatal[comp$fatal == "FATAL"] <- "F"
-```
-
-```{r fatal-eda}
-temp <- comp %>% 
-  group_by(fatal) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=fatal, y=count)) + 
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-```
-```{r race-eda-cities}
-temp <- comp %>% 
-  group_by(fatal, race) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
-```
-
-```{r gender-eda-cities}
-temp <- comp %>% 
-  group_by(fatal, gender) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
-```

### 2020-10-14 submissionii.Rmd
First commit

---
title: "Police Shootings and Racism in America"
author: "Connie Wu"
date: "10/12/2020"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## I. Introduction

|        On May 25, 2020, George Floyd was publicly suffocated to death by police officers in Minneapolis, Minnesota. This, along with the murder of Breonna Taylor by Louisville police, incited waves of protests against police brutality across the United States and increased the spread of Black Lives Matter content on social media. The recent boost in attention to the Black Lives Matter movement has once again brought to light the issue of racism in America and its link to police brutality, and more specifically, police use of deadly force. Several studies have already found that race does play a part in who is targeted and killed in police shootings. For example, in a recent VICE News investigation about police shootings, it was found that “Black people were shot more often and at higher rates than people of any other race.” [1] Additionally, Edwards et. al performed a study regarding the effects of age, race-ethnicity, and sex on the risk of being killed by lethal force by law enforcement and similarly found that "Black men are about 2.5 times more likely to be killed by police over the life course than are white men" while "Black women are about 1.4 times more likely to be killed by police than are white women." [2] However, even with these studies and their disturbing conclusions, a poll done by AP-NORC in June 2020 found that still 39% of Americans think that police are not more likely to use lethal force against a black person than a white person. [3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide even more statistical evidence backing the relationship between racism and police use of deadly force. 
|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings. Additionally, to add predictors relating to location, I will be using data found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs. non-lethal shootings. I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers. Finally, RACE DEMOGRAPHICS SOMEHOW. This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.

### Data
|        To effectively build off of the study done by VICE News, I will be using the same dataset that they used. This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects over seven years.” [1] The predictors that I will use are subject (victim) race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying. The response I will be modeling will be whether the shooting was fatal or not. I will be filtering out the cases in which fatality is unknown or the subject’s or the officer’s race or gender is unknown from the final dataset. Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information. I will carry out the same procedure in the cases in which there were more than one officers present at the shooting. Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.

|        After cleaning the data, I was left with a total of 3,451 observations. I plotted some visualizations that depict the overall trends of the different predictors that I plan to use to answer my proposed research questions.

### Exploratory Data Analysis

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
```

```{r load-data}
all <- read.csv("allshootingsdata.csv")

philly <- read.csv("philly_shootings.csv")
cincinnati <- read.csv("cincinnati_shootings.csv")
```

```{r clean-all}
all <- all %>% 
  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))

all <- all %>% 
  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")

officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
out <- which((officer_gender == officer_race) == FALSE)
all <- all[-out,]

all <- all %>% 
  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
  separate_rows(OfficerRace, OfficerGender, sep = ",")

officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
out2 <- which((officer_gender2 == officer_race2) == FALSE)
all <- all[-out2,]

all$OfficerRace <- trimws(all$OfficerRace)
all$OfficerGender <- trimws(all$OfficerGender)
all$Fatal <- trimws(all$Fatal)

all <- all %>% 
  filter(SubjectRace != "U") %>% 
  filter(OfficerRace != "U") %>% 
  filter(OfficerRace != "Unknown") %>% 
  filter(OfficerRace != "NA") %>% 
  filter(OfficerRace != "") %>% 
  filter(OfficerRace != "I") %>% 
  filter(OfficerGender != "U") %>% 
  filter(OfficerGender != "N") %>% 
  filter(SubjectGender != "U") %>% 
  filter(SubjectGender != "N") %>% 
  filter(Fatal != "U")

all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"

all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
all$OfficerGender[all$OfficerGender == "Male"] <- "M"
```

```{r fatal-eda2}
temp <- all %>% 
  group_by(Fatal) %>% 
  summarise(count = n())

ggplot(temp, aes(x=Fatal, y=count)) + 
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
```

|        In this dataset, it is important to keep in mind that there is a higher number of non-fatal shooting events than fatal shooting events.

```{r race-eda}
temp <- all %>% 
  group_by(Fatal, SubjectRace) %>% 
  summarise(race_count = n())

ggplot(temp, aes(x=SubjectRace, y=race_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in Police Shootings")

temp2 <- all %>% 
  group_by(Fatal, OfficerRace) %>% 
  summarise(race_count = n())

ggplot(temp2, aes(x=OfficerRace, y=race_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
```

|        In the plots above, we can see the counts for both victim and officer race in the final dataset split by fatal versus non-fatal shootings. In this dataset, A stands for Asian, B stands for Black, H stands for Hispanic, L stands for Latino, O stands for other, and W stands for White. Already, we can see that in both fatal and non-fatal cases, Black is the race with the highest count for the victim category, whereas White is the race with the highest count for the officer category. It is also interesting to note that the number of Black victims is higher in non-fatal cases than in fatal cases, while the number of white victims is higher in fatal cases than in non-fatal cases.

```{r gender-eda}
temp <- all %>% 
  group_by(Fatal, SubjectGender) %>% 
  summarise(gender_count = n())

ggplot(temp, aes(x=SubjectGender, y=gender_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in Police Shootings")

temp2 <- all %>% 
  group_by(Fatal, OfficerGender) %>% 
  summarise(gender_count = n())

ggplot(temp2, aes(x=OfficerGender, y=gender_count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
```

|        Additionally, it is evident from the graphs above that male is the gender with the highest count in the cases of victims and officers for both fatal and non-fatal shootings.

```{r weapon-eda}
temp <- all %>% 
  group_by(Fatal, weapon) %>% 
  summarise(count = n())

ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings") 
```

|        Finally, we can see that in cases of police shootings, victims have guns the most often in fatal cases and unknown weapons in non-fatal cases, whereas they have gun replicas the least often in both cases.

|        As I mentioned above, in order to draw comparisons between the demographics of those involved in police shootings versus non-police shootings, I will be using data collected by the Police Data Initiative on the shootings in two cities (Philadelphia and Cincinnati). I will only be using data from these two cities because obtaining data on both fatal and non-fatal shooting crime in America is extremely difficult, since non-fatal shootings are “generally lumped-in with other types of assault in national data collections.” [2] The Police Data Initiative also suggests using general shooting data from Rochester, New York. However, after looking into this dataset, I found that it did not specify whether each shooting was fatal or non-fatal. As a result, I have decided to leave Rochester out of my final dataset so that I can primarily focus on identifying trends in victim and officer demographics in the context of fatal versus non-fatal shootings. The main predictors I will use from the Philadelphia and Cincinnati datasets will be race and sex, both of which refer to the victims of the shootings. I will combine these selected variables from the two datasets into one final dataset. Once again, I will be using fatality of the shooting as the response variable.

```{r clean-cities}
cinn <- cincinnati %>% 
  select(RACE, SEX, TYPE) %>% 
  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
  mutate(city = "Cincinnati")

phil <- philly %>% 
  select(race, sex, fatal) %>% 
  rename(gender = sex) %>% 
  mutate(city = "Philadelphia")

comp <- rbind(cinn, phil)

comp <- comp %>%
  filter(race != "UNKNOWN") %>% 
  filter (race != "") %>% 
  filter(race != "UNK") %>% 
  filter(gender != "")

comp$race[comp$race == "WHITE"] <- "W"
comp$race[comp$race == "BLACK"] <- "B"
comp$race[comp$race == "OTHER"] <- "O"
comp$race[comp$race == "HISPANIC"] <- "H"

comp$gender[comp$gender == "MALE"] <- "M"
comp$gender[comp$gender == "FEMALE"] <- "F"

comp$fatal[comp$fatal == "0"] <- "N"
comp$fatal[comp$fatal == "1"] <- "F"
comp$fatal[comp$fatal == "NONFATAL"] <- "N"
comp$fatal[comp$fatal == "FATAL"] <- "F"
```

```{r fatal-eda1}
temp <- comp %>% 
  group_by(fatal) %>% 
  summarise(count = n())

ggplot(temp, aes(x=fatal, y=count)) + 
  geom_bar(stat="identity") + 
  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
```

|        In this dataset, it is important to keep in mind that again, there is a significantly higher number of non-fatal shooting events than fatal shooting events.

```{r race-eda-cities}
temp <- comp %>% 
  group_by(fatal, race) %>% 
  summarise(count = n())

ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
```

```{r gender-eda-cities}
temp <- comp %>% 
  group_by(fatal, gender) %>% 
  summarise(count = n())

ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
  geom_bar(position="dodge", stat="identity") + 
  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
```

|        Furthermore, it appears that Black is the race with the highest count and male is the gender with the highest count for victims of all shootings in these two cities, which is similar to what we discovered for the police shooting data. However, it does not seem that this trend in race and gender varies too much between fatal and non-fatal shootings.

## II. Methods

### Data Processing

### Model 
```{r}

```

## III. Preliminary Results

## References 
1) Vice 
2) https://www.pnas.org/content/pnas/116/34/16793.full.pdf
3) https://www.kpbs.org/news/2020/jun/18/sweeping-change-us-views-police-violence-new-poll-/
4) http://govred.com/blog/deescalation-training-state-requirements/

### 2020-10-15 submissionii.Rmd
data added

@@
 al performed a study regarding the effects of age, race-ethnicity, and sex on the risk of being killed by lethal force by law enforcement and similarly found that "Black men are about 2.5 times more likely to be killed by police over the life course than are white men" while "Black women are about 1.4 times more likely to be killed by police than are white women." [2] However, even with these studies and their disturbing conclusions, a poll done by AP-NORC in June 2020 found that still 39% of Americans think that police are not more likely to use lethal force against a black person than a white person.
-[3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide even more statistical evidence backing the relationship between racism and police use of deadly force.
-|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
-Additionally, to add predictors relating to location, I will be using data found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
+[3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
+|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
+Additionally, to account for the varying locations of the homicides, I will be using data found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
 non-lethal shootings.
 I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers.
-Finally, RACE DEMOGRAPHICS SOMEHOW.
+Finally, I will explore how the ratio of black to white population in each location relates to police use of deadly force.
 This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.
@@
 ### Data
-|        To effectively build off of the study done by VICE News, I will be using the same dataset that they used.
+|        As a basis for this study, I will be using the same dataset that VICE News used.
 This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects over seven years.” [1] The predictors that I will use are subject (victim) race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
@@
 ```{r load-data}
-all <- read.csv("allshootingsdata.csv")
-
-philly <- read.csv("philly_shootings.csv")
-cincinnati <- read.csv("cincinnati_shootings.csv")
+all <- read.csv("data/allshootingsdata.csv")
+census <- read.csv("data/2013census.csv")[,-c(8,9)]
+gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
+  filter(year == 2013)
 ```

### 2020-10-16 submissionii.Rmd
data manipulated kinda

@@
 ```{r load-data}
+# Shootings
 all <- read.csv("data/allshootingsdata.csv")
+all <- all %>% 
+  mutate(id = paste0("C", row_number()))
+
+# Census Predictors
 census <- read.csv("data/2013census.csv")[,-c(8,9)]
+states <- lapply(census$department, function(x) str_extract_all(x,"\\([^()]+\\)")[[1]])
+states <- substring(states, 2, nchar(states)-1)
+census$state <- states
+
+# Gun Law Predictors
 gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
   filter(year == 2013)
+gun_provisions <- gun_provisions %>% 
+  mutate(total_pr = rowSums(gun_provisions[,3:136]),
+         deesc_train = ifelse(state %in% c("California", "New Mexico", "Texas", "Oklahoma", "South Dakota", "Illinois", "Indiana", "Tennessee", "Washington", "Missouri", "Georgia", "Ohio", "Maryland", "Massachusetts", "Connecticut", "Maine"), 1, 0)) %>%
+  select(state, total_pr, deesc_train)
+gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
 ```

### 2020-10-16 submissionii.Rmd
thinking bout model...

@@
 all <- all %>% 
-  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))
+  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit"))
 

### 2020-10-16 police_modeling.Rmd
added to data section

@@
 ```{r load-data}
-all <- read.csv("allshootingsdata.csv")
+# Shootings
+all <- read.csv("data/allshootingsdata.csv")
+all <- all %>% 
+  mutate(id = paste0("C", row_number()))
 
-philly <- read.csv("philly_shootings.csv")
-cincinnati <- read.csv("cincinnati_shootings.csv")
+# Census Predictors
+census <- read.csv("data/2013census.csv")[,-c(8,9)]
+states <- lapply(census$department, function(x) str_extract_all(x,"\\([^()]+\\)")[[1]])
+states <- substring(states, 2, nchar(states)-1)
+census$state <- states
+
+# Gun Law Predictors
+gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
+  filter(year == 2013)
+gun_provisions <- gun_provisions %>% 
+  mutate(total_pr = rowSums(gun_provisions[,3:136]),
+         deesc_train = ifelse(state %in% c("California", "New Mexico", "Texas", "Oklahoma", "South Dakota", "Illinois", "Indiana", "Tennessee", "Washington", "Missouri", "Georgia", "Ohio", "Maryland", "Massachusetts", "Connecticut", "Maine"), 1, 0)) %>%
+  select(state, total_pr, deesc_train)
+gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
 ```

### 2020-10-16 submissionii.Rmd
added to data section

@@
 
-## I. Introduction
-
-|        On May 25, 2020, George Floyd was publicly suffocated to death by police officers in Minneapolis, Minnesota.
-This, along with the murder of Breonna Taylor by Louisville police, incited waves of protests against police brutality across the United States and increased the spread of Black Lives Matter content on social media.
-The recent boost in attention to the Black Lives Matter movement has once again brought to light the issue of racism in America and its link to police brutality, and more specifically, police use of deadly force.
-Several studies have already found that race does play a part in who is targeted and killed in police shootings.
-For example, in a recent VICE News investigation about police shootings, it was found that “Black people were shot more often and at higher rates than people of any other race.” [1] Additionally, Edwards et.
-al performed a study regarding the effects of age, race-ethnicity, and sex on the risk of being killed by lethal force by law enforcement and similarly found that "Black men are about 2.5 times more likely to be killed by police over the life course than are white men" while "Black women are about 1.4 times more likely to be killed by police than are white women." [2] However, even with these studies and their disturbing conclusions, a poll done by AP-NORC in June 2020 found that still 39% of Americans think that police are not more likely to use lethal force against a black person than a white person.
-[3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
-|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
-Additionally, to account for the varying locations of the homicides, I will be using data found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
-non-lethal shootings.
-I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers.
-Finally, I will explore how the ratio of black to white population in each location relates to police use of deadly force.
-This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.
-
-### Data
-|        As a basis for this study, I will be using the same dataset that VICE News used.
-This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects over seven years.” [1] The predictors that I will use are subject (victim) race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
-The response I will be modeling will be whether the shooting was fatal or not.
-I will be filtering out the cases in which fatality is unknown or the subject’s or the officer’s race or gender is unknown from the final dataset.
-Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information.
-I will carry out the same procedure in the cases in which there were more than one officers present at the shooting.
-Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.
-
-|        After cleaning the data, I was left with a total of 3,451 observations.
-I plotted some visualizations that depict the overall trends of the different predictors that I plan to use to answer my proposed research questions.
-
-### Exploratory Data Analysis
-
 ```{r}
@@
 all <- read.csv("data/allshootingsdata.csv")
-all <- all %>% 
-  mutate(id = paste0("C", row_number()))
 
@@
 all <- all %>% 
-  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit"))
+  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
+  mutate(other = as.numeric(other), 
+         replica = as.numeric(replica), 
+         knife = as.numeric(knife), 
+         gun = as.numeric(weapon == "gun"), 
+         unarmed = as.numeric(weapon == "unarmed"),
+         unknown_weapon = as.numeric(weapon == "U"),
+         id = paste0("C", row_number())) %>% 
+  select(-weapon)
 
@@
 ```
+
+## I. Introduction
+
+|        On May 25, 2020, George Floyd was publicly suffocated to death by police officers in Minneapolis, Minnesota.
+This, along with the murder of Breonna Taylor by Louisville police, incited waves of protests against police brutality across the United States and increased the spread of Black Lives Matter content on social media.
+The recent boost in attention to the Black Lives Matter movement has once again brought to light the issue of racism in America and its link to police brutality, and more specifically, police use of deadly force.
+Several studies have already found that race does play a part in who is targeted and killed in police shootings.
+For example, in a recent VICE News investigation about police shootings, it was found that “Black people were shot more often and at higher rates than people of any other race.” [1] Additionally, Edwards et.
+al performed a study regarding the effects of age, race-ethnicity, and sex on the risk of being killed by lethal force by law enforcement and similarly found that "Black men are about 2.5 times more likely to be killed by police over the life course than are white men" while "Black women are about 1.4 times more likely to be killed by police than are white women." [2] However, even with these studies and their disturbing conclusions, a poll done by AP-NORC in June 2020 found that still 39% of Americans think that police are not more likely to use lethal force against a black person than a white person.
+[3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
+|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
+Additionally, to account for the varying locations of the homicides, I will be using data found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
+non-lethal shootings.
+I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers.
+Finally, I will explore how the ratio of black to white population in each location relates to police use of deadly force.
+This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.
+
+### Data
+|        As a basis for this study, I will be using the same dataset that VICE News used.
+This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including date, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
+Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), and Unknown (U).
+Gender was broken up into 3 categories: Male (M), Female (F), and Unknown (U).
+Additionally, there were cases in which multiple victims and/or multiple officers were present in the shooting.
+Each of these scenarios was still represented within one row of the dataset, but semi-colons were present in the respective victim and officer columns, separating each individual's information from each other.
+
+### Exploratory Data Analysis
 
@@
 
-```{r weapon-eda}
+```{r weapon-eda, eval=F}
 temp <- all %>% 
@@
   summarise(count = n())
-
 ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
@@
 
-|        As I mentioned above, in order to draw comparisons between the demographics of those involved in police shootings versus non-police shootings, I will be using data collected by the Police Data Initiative on the shootings in two cities (Philadelphia and Cincinnati).
-I will only be using data from these two cities because obtaining data on both fatal and non-fatal shooting crime in America is extremely difficult, since non-fatal shootings are “generally lumped-in with other types of assault in national data collections.” [2] The Police Data Initiative also suggests using general shooting data from Rochester, New York.
-However, after looking into this dataset, I found that it did not specify whether each shooting was fatal or non-fatal.
-As a result, I have decided to leave Rochester out of my final dataset so that I can primarily focus on identifying trends in victim and officer demographics in the context of fatal versus non-fatal shootings.
-The main predictors I will use from the Philadelphia and Cincinnati datasets will be race and sex, both of which refer to the victims of the shootings.
-I will combine these selected variables from the two datasets into one final dataset.
-Once again, I will be using fatality of the shooting as the response variable.
-
-```{r clean-cities}
-cinn <- cincinnati %>% 
-  select(RACE, SEX, TYPE) %>% 
-  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
-  mutate(city = "Cincinnati")
-
-phil <- philly %>% 
-  select(race, sex, fatal) %>% 
-  rename(gender = sex) %>% 
-  mutate(city = "Philadelphia")
-
-comp <- rbind(cinn, phil)
-
-comp <- comp %>%
-  filter(race != "UNKNOWN") %>% 
-  filter (race != "") %>% 
-  filter(race != "UNK") %>% 
-  filter(gender != "")
-
-comp$race[comp$race == "WHITE"] <- "W"
-comp$race[comp$race == "BLACK"] <- "B"
-comp$race[comp$race == "OTHER"] <- "O"
-comp$race[comp$race == "HISPANIC"] <- "H"
-
-comp$gender[comp$gender == "MALE"] <- "M"
-comp$gender[comp$gender == "FEMALE"] <- "F"
-
-comp$fatal[comp$fatal == "0"] <- "N"
-comp$fatal[comp$fatal == "1"] <- "F"
-comp$fatal[comp$fatal == "NONFATAL"] <- "N"
-comp$fatal[comp$fatal == "FATAL"] <- "F"
-```
-
-```{r fatal-eda1}
-temp <- comp %>% 
-  group_by(fatal) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=fatal, y=count)) + 
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-```
-
-|        In this dataset, it is important to keep in mind that again, there is a significantly higher number of non-fatal shooting events than fatal shooting events.
-
-```{r race-eda-cities}
-temp <- comp %>% 
-  group_by(fatal, race) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
-```
-
-```{r gender-eda-cities}
-temp <- comp %>% 
-  group_by(fatal, gender) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
-```
-
-|        Furthermore, it appears that Black is the race with the highest count and male is the gender with the highest count for victims of all shootings in these two cities, which is similar to what we discovered for the police shooting data.
-However, it does not seem that this trend in race and gender varies too much between fatal and non-fatal shootings.
-
 ## II. Methods
@@
 
+|        In the VICE dataset, I will be filtering out the cases in which fatality is unknown, or the subject’s or officer’s race or gender is unknown from the final dataset.
+Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information.
+I will carry out the same procedure in the cases in which there were more than one officers present at the shooting.
+Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.
+
+|        After cleaning the data, I was left with a total of 3,451 observations.
+I plotted some visualizations that depict the overall trends of the different predictors that I plan to use to answer my proposed research questions.
+
 ### Model 
+The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
+The response I will be modeling will be whether the shooting was fatal or not.
 ```{r}
-
+fuck
 ```

### 2020-10-16 submissionii.Rmd
added to data section
finished data section

@@
 [3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
+
 |        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
-Additionally, to account for the varying locations of the homicides, I will be using data found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
+Additionally, to account for the varying locations of the homicides, I will be using a dataset found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
 non-lethal shootings.
@@
 ### Data
+
 |        As a basis for this study, I will be using the same dataset that VICE News used.
-This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including date, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
+This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from 2010 to 2016, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
 Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), and Unknown (U).
@@
 Each of these scenarios was still represented within one row of the dataset, but semi-colons were present in the respective victim and officer columns, separating each individual's information from each other.
+
+|       Additionally, to assess the effect of different types of legislation on fatal vs.
+non-fatal police shootings in various states, I will be using a dataset from Kaggle containing 135 variables that detail whether a certain gun provision is absent or present in a certain year and U.S.
+state for 133 different gun provisions.
+The years range from 1991 to 2017, and the gun provisions address 14 categories, some of which are dealer regulations, buyer regulations, prohibitions for high-risk gun possession, background checks, ammunition regulations, possession regulations, concealed carry permitting, assault weapons and large-capacity magazines, child access prevention, gun trafficking, and domestic violence.
+A 1 in the gun provision column represents a presence of the law, and a 0 indicates an absence.
+I also will add a column to this gun provision dataset indicating whether a state requires de-escalation training based off of information reported by APM Report in 2017, with a 1 meaning the state does require training and a 0 meaning the state does not.[4]
+
+|       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets.
+VICE News most likely provided ony 2013 census data because it is the average year of all the years represented in the shootings dataset.
+The census data includes 7 variables: the police department and the city it's located in, as well as the city's black, asian, hispanic, white, and overall total population in 2013.
 
@@
 ```{r}
-fuck
+
 ```

### 2020-10-17 police_eda.Rmd
cleaning office race

@@
 ## Load Data
-```{r load-data}
-all <- read.csv("allshootingsdata.csv")
 
-philly <- read.csv("philly_shootings.csv")
-cincinnati <- read.csv("cincinnati_shootings.csv")
+```{r}
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+library(tidyverse)
 ```
 
-## Exploratory Data Analysis
+```{r load-data}
+# Shootings
+all <- read.csv("data/allshootingsdata.csv")
+
+# Census Predictors
+census <- read.csv("data/2013census.csv")[,-c(8,9)]
+states <- lapply(census$department, function(x) str_extract_all(x,"\\([^()]+\\)")[[1]])
+states <- substring(states, 2, nchar(states)-1)
+census$state <- states
+
+# Gun Law Predictors
+gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
+  filter(year == 2013)
+gun_provisions <- gun_provisions %>% 
+  mutate(total_pr = rowSums(gun_provisions[,3:136]),
+         deesc_train = ifelse(state %in% c("California", "New Mexico", "Texas", "Oklahoma", "South Dakota", "Illinois", "Indiana", "Tennessee", "Washington", "Missouri", "Georgia", "Ohio", "Maryland", "Massachusetts", "Connecticut", "Maine"), 1, 0)) %>%
+  select(state, total_pr, deesc_train)
+gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
+```
 
@@
 all <- all %>% 
-  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))
+  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
+  mutate(other = as.numeric(other), 
+         replica = as.numeric(replica), 
+         knife = as.numeric(knife), 
+         gun = as.numeric(weapon == "gun"), 
+         unarmed = as.numeric(weapon == "unarmed"),
+         unknown_weapon = as.numeric(weapon == "U"),
+         id = paste0("C", row_number())) %>% 
+  select(-weapon)
 
@@
 all$OfficerGender[all$OfficerGender == "Male"] <- "M"
+```
+
+```{r split_data}
+single_victims <- all %>% 
+  filter(!str_detect(SubjectRace, ";"))
+
+mult_victims <- all %>% 
+  filter(str_detect(SubjectRace, ";"))
+```
+
+## Exploratory Data Analysis
+
+```{r}
+temp <- single_victims %>% 
+  group_by(Fatal, SubjectRace) %>% 
+  summarise(count = n())
+
+ggplot(single_victims, aes(x = Fatal, y = count)) +
+  geom_bar(stat="identity") + 
+  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
+  group_by(Fatal, race) %>% 
+  summarise(count = n())
 ```
@@
 ```
-
-```{r clean-cities}
-cinn <- cincinnati %>% 
-  select(RACE, SEX, TYPE) %>% 
-  rename(race = RACE, gender = SEX, fatal = TYPE) %>% 
-  mutate(city = "Cincinnati")
-
-phil <- philly %>% 
-  select(race, sex, fatal) %>% 
-  rename(gender = sex) %>% 
-  mutate(city = "Philadelphia")
-
-comp <- rbind(cinn, phil)
-
-comp <- comp %>%
-  filter(race != "UNKNOWN") %>% 
-  filter (race != "") %>% 
-  filter(race != "UNK") %>% 
-  filter(gender != "")
-
-comp$race[comp$race == "WHITE"] <- "W"
-comp$race[comp$race == "BLACK"] <- "B"
-comp$race[comp$race == "OTHER"] <- "O"
-comp$race[comp$race == "HISPANIC"] <- "H"
-
-comp$gender[comp$gender == "MALE"] <- "M"
-comp$gender[comp$gender == "FEMALE"] <- "F"
-
-comp$fatal[comp$fatal == "0"] <- "N"
-comp$fatal[comp$fatal == "1"] <- "F"
-comp$fatal[comp$fatal == "NONFATAL"] <- "N"
-comp$fatal[comp$fatal == "FATAL"] <- "F"
-```
-
-```{r fatal-eda}
-temp <- comp %>% 
-  group_by(fatal) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=fatal, y=count)) + 
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-```
-```{r race-eda-cities}
-temp <- comp %>% 
-  group_by(fatal, race) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=race, y=count, fill = fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in All Shootings") 
-```
-
-```{r gender-eda-cities}
-temp <- comp %>% 
-  group_by(fatal, gender) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=gender, y=count, fill = fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in All Shootings") 
-```

### 2020-10-17 submissionii.Rmd
cleaning office race

@@
 ```{r clean-all}
+all$OfficerRace <- gsub("H/L", "L", all$OfficerRace)
+all$OfficerRace <- gsub("H", "L", all$OfficerRace)
+all$OfficerRace <- gsub("Unknown", "U", all$OfficerRace)
+all$OfficerRace <- gsub("m/m", "U", all$OfficerRace)
+all$OfficerRace <- gsub("WHITE", "W", all$OfficerRace)
+all$OfficerRace <- gsub("WLITE", "W", all$OfficerRace)
+all$OfficerRace <- gsub("BLACK", "B", all$OfficerRace)
+all$OfficerRace <- gsub("ASIAN", "A", all$OfficerRace)
+all$OfficerRace <- gsub("Other", "O", all$OfficerRace)
+all$OfficerRace <- gsub("AI/AN", "O", all$OfficerRace)
+all$OfficerRace <- gsub("A/PI ", "A;", all$OfficerRace)
+all$OfficerRace <- gsub("A/PI", "A", all$OfficerRace)
+all$OfficerRace <- gsub("Multi-Racial", "O", all$OfficerRace)
+all$OfficerRace <- gsub(",", ";", all$OfficerRace)
+all$OfficerRace <- gsub(" ", "", all$OfficerRace)
+all$OfficerRace <- gsub("A/W", "O", all$OfficerRace)
+all$OfficerRace <- gsub("NA/W", "O", all$OfficerRace)
+all$OfficerRace <- gsub("W/L", "O", all$OfficerRace)
+all$OfficerRace <- gsub("W/A", "O", all$OfficerRace)
+
 all <- all %>% 
   select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
-  mutate(other = as.numeric(other), 
+  mutate(b = as.numeric(b > 0),
+         w = as.numeric(w > 0),
+         l = as.numeric(l > 0),
+         a = as.numeric(a > 0),
+         other = as.numeric(other), 
          replica = as.numeric(replica), 
@@
   select(-weapon)
+
+all <- rename(all, sb = b, sl = l, sw = w, sa = a)
+
+all$ob <- as.numeric(str_count(all$OfficerRace, "B") > 0)
+all$ow <- as.numeric(str_count(all$OfficerRace, "W") > 0)
+all$ol <- as.numeric(str_count(all$OfficerRace, "L") > 0)
+all$oa <- as.numeric(str_count(all$OfficerRace, "A") > 0)
 
@@
 
-|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
+|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders, whether they are carrying a weapon, and the total number of victims in the crime, play in fatal versus non-fatal police shootings.
 Additionally, to account for the varying locations of the homicides, I will be using a dataset found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
@@
 |       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets.
-VICE News most likely provided ony 2013 census data because it is the average year of all the years represented in the shootings dataset.
-The census data includes 7 variables: the police department and the city it's located in, as well as the city's black, asian, hispanic, white, and overall total population in 2013.
+VICE News most likely provided only 2013 census data because it is the average year of all the years represented in the shootings dataset.
+The census data includes 7 variables: the police department and the city it's located in, as well as the city's Black, Asian, Hispanic, White, and overall total population in 2013.
 
 ### Exploratory Data Analysis
+
+```{r}
+single_victims <- all %>% 
+  filter(!str_detect(SubjectRace, ";"))
+nrow(single_victims)
+
+mult_victims <- all %>% 
+  filter(str_detect(SubjectRace, ";"))
+nrow(mult_victims)
+
+unknown <- all %>% 
+  filter(str_detect(SubjectRace, "U"))
+```
+|       Upon first glance, there appears to be 3906 incidents in which there was only one victim, and 210 with multiple victims.
+This means that there is one incident in which the victim race was marked as "NA." Additionally,
+
+```{r}
+single_victims <- all %>% 
+  filter(!str_detect(SubjectRace, ";"))
+
+mult_victims <- all %>% 
+  filter(str_detect(SubjectRace, ";"))
+
+ggplot(single_victims, aes(x = Fatal, y = count)) +
+  geom_bar(stat="identity") + 
+  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
+  group_by(Fatal, race) %>% 
+  summarise(count = n())
+```
 

### 2020-10-17 police_eda.Rmd
added some data processing

@@
 ```{r load-data}
-# Shootings
-all <- read.csv("data/allshootingsdata.csv")
-
 # Census Predictors
@@
 
+### Clean all
+
 ```{r clean-all}
+all <- read.csv("data/allshootingsdata.csv")
+
+# Recode subject race
+all$SubjectRace[is.na(all$SubjectRace)] <- "U"
+
+# Recode officer races
+all$OfficerRace <- gsub("H/L", "L", all$OfficerRace)
+all$OfficerRace <- gsub("H", "L", all$OfficerRace)
+all$OfficerRace <- gsub("Unknown|m/m", "U", all$OfficerRace)
+all$OfficerRace <- gsub("WHITE|WLITE", "W", all$OfficerRace)
+all$OfficerRace <- gsub("BLACK", "B", all$OfficerRace)
+all$OfficerRace <- gsub("ASIAN", "A", all$OfficerRace)
+all$OfficerRace <- gsub("Other|AI/AN|NA/W|Multi-Racial", "O", all$OfficerRace)
+all$OfficerRace <- gsub("A/PI ", "A;", all$OfficerRace)
+all$OfficerRace <- gsub("A/PI", "A", all$OfficerRace)
+all$OfficerRace <- gsub(",", ";", all$OfficerRace)
+all$OfficerRace <- gsub(" ", "", all$OfficerRace)
+all$OfficerRace <- gsub("A/W|W/L|W/A", "O", all$OfficerRace)
+all$OfficerRace[all$OfficerRace == ""] <- "U"
+
+# Recode officer genders
+all$OfficerGender <- gsub("MALE", "M", toupper(all$OfficerGender))
+all$OfficerGender <- gsub("FEMALE|FEM", "F", toupper(all$OfficerGender))
+all$OfficerGender <- gsub("UNKNOWN|FALSE", "U", toupper(all$OfficerGender))
+all$OfficerGender <- gsub(",|:", ";", toupper(all$OfficerGender))
+all$OfficerGender <- gsub(" |/", "", toupper(all$OfficerGender))
+
 all <- all %>% 
   select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
-  mutate(other = as.numeric(other), 
+  mutate(b = as.numeric(b > 0),
+         w = as.numeric(w > 0),
+         l = as.numeric(l > 0),
+         a = as.numeric(a > 0),
+         so =  as.numeric(str_count(all$SubjectRace, "O") > 0),
+         smale = as.numeric(str_count(all$SubjectGender, "M") > 0),
+         sfem = as.numeric(str_count(all$SubjectGender, "F") > 0),
+         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
+         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
+         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
+         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
+         oo =  as.numeric(str_count(all$OfficerRace, "O") > 0),
+         omale = as.numeric(str_count(all$OfficerGender, "M") > 0),
+         ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
+         other = as.numeric(other), 
          replica = as.numeric(replica), 
@@
          unarmed = as.numeric(weapon == "unarmed"),
-         unknown_weapon = as.numeric(weapon == "U"),
-         id = paste0("C", row_number())) %>% 
-  select(-weapon)
+         unknown_weapon = as.numeric(weapon == "U")) %>% 
+  select(-c(weapon, SubjectRace, SubjectGender, OfficerRace, OfficerGender))
 
+all <- rename(all, sb = b, sw = w, sl = l, sa = a)
+```
+
+```{r filter-all}
+# Filter out unknown values
+all <- all[-which(all$smale == 0 & all$sfem == 0),]
+all <- all[-which(all$omale == 0 & all$ofem == 0),]
+all <- all[-which(all$sb == 0 & all$sw == 0 & all$sl == 0 & all$sa == 0),]
+all <- all[-which(all$ob == 0 & all$ow == 0 & all$ol == 0 & all$oa == 0),]
+```
+
+```{r save-all}
+all <- saveRDS(all, "data/final.rds")
+```
+
+```{r}
 all <- all %>% 
@@
 all$OfficerGender[all$OfficerGender == "Male"] <- "M"
+
 ```

### 2020-10-17 submissionii.Rmd
added some data processing

--- 2020-10-16 submissionii.Rmd
@@
 gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
+
+# Shootings
+all <- read.csv("data/allshootingsdata.csv")
 ```
@@
 ```{r clean-all}
+# Recode officer races
+all$OfficerRace <- gsub("H/L", "L", all$OfficerRace)
+all$OfficerRace <- gsub("H", "L", all$OfficerRace)
+all$OfficerRace <- gsub("Unknown|m/m", "U", all$OfficerRace)
+all$OfficerRace <- gsub("WHITE|WLITE", "W", all$OfficerRace)
+all$OfficerRace <- gsub("BLACK", "B", all$OfficerRace)
+all$OfficerRace <- gsub("ASIAN", "A", all$OfficerRace)
+all$OfficerRace <- gsub("Other|AI/AN|NA/W|Multi-Racial", "O", all$OfficerRace)
+all$OfficerRace <- gsub("A/PI ", "A;", all$OfficerRace)
+all$OfficerRace <- gsub("A/PI", "A", all$OfficerRace)
+all$OfficerRace <- gsub(",", ";", all$OfficerRace)
+all$OfficerRace <- gsub(" ", "", all$OfficerRace)
+all$OfficerRace <- gsub("A/W|W/L|W/A", "O", all$OfficerRace)
+all$OfficerRace[all$OfficerRace == ""] <- "U"
+
+# Recode officer genders
+all$OfficerGender <- gsub("MALE", "M", toupper(all$OfficerGender))
+all$OfficerGender <- gsub("FEMALE|FEM", "F", toupper(all$OfficerGender))
+all$OfficerGender <- gsub("UNKNOWN|FALSE", "U", toupper(all$OfficerGender))
+all$OfficerGender <- gsub(",|:", ";", toupper(all$OfficerGender))
+all$OfficerGender <- gsub(" |/", "", toupper(all$OfficerGender))
+
 all <- all %>% 
   select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
-  mutate(other = as.numeric(other), 
+  mutate(b = as.numeric(b > 0),
+         w = as.numeric(w > 0),
+         l = as.numeric(l > 0),
+         a = as.numeric(a > 0),
+         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
+         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
+         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
+         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
+         other = as.numeric(other), 
          replica = as.numeric(replica), 
@@
          id = paste0("C", row_number())) %>% 
-  select(-weapon)
+  select(-c(weapon))
+
+all <- rename(all, sb = b, sl = l, sw = w, sa = a)
 
@@
 For example, in a recent VICE News investigation about police shootings, it was found that “Black people were shot more often and at higher rates than people of any other race.” [1] Additionally, Edwards et.
-al performed a study regarding the effects of age, race-ethnicity, and sex on the risk of being killed by lethal force by law enforcement and similarly found that "Black men are about 2.5 times more likely to be killed by police over the life course than are white men" while "Black women are about 1.4 times more likely to be killed by police than are white women." [2] However, even with these studies and their disturbing conclusions, a poll done by AP-NORC in June 2020 found that still 39% of Americans think that police are not more likely to use lethal force against a black person than a white person.
+al performed a study regarding the effects of age, race-ethnicity, and sex on the risk of being killed by lethal force by law enforcement and similarly found that "Black men are about 2.5 times more likely to be killed by police over the life course than are white men" while "Black women are about 1.4 times more likely to be killed by police than are white women." [2] However, even with these studies and their disturbing conclusions, a poll done by AP-NORC in June 2020 found that still 39% of Americans think that police are not more likely to use lethal force against a Black person than a White person.
 [3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
 
-|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders and whether they are carrying a weapon, play in fatal versus non-fatal police shootings.
+|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders, whether they are carrying a weapon, and the total number of victims in the crime, play in fatal versus non-fatal police shootings.
 Additionally, to account for the varying locations of the homicides, I will be using a dataset found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
@@
 I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers.
-Finally, I will explore how the ratio of black to white population in each location relates to police use of deadly force.
+Finally, I will explore how the ratio of Black to White population in each location relates to police use of deadly force.
 This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.
@@
 This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from 2010 to 2016, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
-Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), and Unknown (U).
+Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), Other (O), and Unknown (U).
 Gender was broken up into 3 categories: Male (M), Female (F), and Unknown (U).
+Weapon type was broken up into 6 values: "gun", "knife", "replica", "other", "unknown", and "other".
 Additionally, there were cases in which multiple victims and/or multiple officers were present in the shooting.
@@
 |       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets.
-VICE News most likely provided ony 2013 census data because it is the average year of all the years represented in the shootings dataset.
-The census data includes 7 variables: the police department and the city it's located in, as well as the city's black, asian, hispanic, white, and overall total population in 2013.
+VICE News most likely provided only 2013 census data because it is the average year of all the years represented in the shootings dataset.
+The census data includes 7 variables: the police department and the city it's located in, as well as the city's Black, Asian, Hispanic, White, and overall total population in 2013.
+
+## II. Methods
+
+### Data Processing
+
+|        Although the VICE News dataset was already relatively clean with regards to the victim's data, there was still quite a bit of cleaning that needed to be done for the race and gender data of the officers.
+To clean these columns, I replaced all values that were not "W", "B", "A", "O", or "U" in the race column with the most informed guess that I could make about what the values represented.
+For example, "A/PI" values were taken to represent Asian/Pacific Islander and thus were replaced with "A", and "A/W" or other values with a "/" in them were typically taken to represent multi-racial individuals and as a result were replaced with "O." Similar procedures were carried out in the gender column for officers for values that were not "M", "F", or "U".
+
+|        After this initial cleaning, I created new columns in the same dataset that represented whether any victims of race Black, White, Asian, or Other were present in the crime, respectively, and if any male or female victims were present in the crime, respectively.
+I also added new columns to represent the same information for officers (for each race and gender, whether each was present).
+I then made sure that each weapon type had its own column.
+All of these added columns had values of either 0 or 1, with 0 representing an absence of the variable and 1 a presence.
+Finally, I filtered out all rows that had only unknown values for the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether the shooting was fatal or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
+
+In the VICE dataset, I will be filtering out the cases in which fatality is unknown, or the subject’s or officer’s race or gender is unknown from the final dataset.
+Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information.
+I will carry out the same procedure in the cases in which there were more than one officers present at the shooting.
+Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.
+
+|        After cleaning the data, I was left with a total of 2,084 observations.
 
 ### Exploratory Data Analysis
+
+```{r}
+single_victims <- all %>% 
+  filter(NumberOfSubjects <= 1)
+nrow(single_victims)
+
+mult_victims <- all %>% 
+  filter(NumberOfSubjects > 1)
+nrow(mult_victims)
+
+unknown <- all %>% 
+  filter(str_detect(SubjectRace, "U"))
+```
+|       Upon first glance, there appears to be 1932 incidents in which there was only one victim, and 152 with multiple victims.
+This means that there is one incident in which the victim race was marked as "NA." Additionally,
+
+```{r eval = F}
+single_victims <- all %>% 
+  filter(!str_detect(SubjectRace, ";"))
+
+mult_victims <- all %>% 
+  filter(str_detect(SubjectRace, ";"))
+
+ggplot(single_victims, aes(x = Fatal, y = count)) +
+  geom_bar(stat="identity") + 
+  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
+  group_by(Fatal, race) %>% 
+  summarise(count = n())
+```
 
@@
 
-## II. Methods
-
-### Data Processing
-
-|        In the VICE dataset, I will be filtering out the cases in which fatality is unknown, or the subject’s or officer’s race or gender is unknown from the final dataset.
-Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information.
-I will carry out the same procedure in the cases in which there were more than one officers present at the shooting.
-Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.
-
-|        After cleaning the data, I was left with a total of 3,451 observations.
-I plotted some visualizations that depict the overall trends of the different predictors that I plan to use to answer my proposed research questions.
-
 ### Model 

### 2020-10-17 police_eda.Rmd
merging dfs

@@
 
-```{r save-all}
-all <- saveRDS(all, "data/final.rds")
+```{r}
+final_df <- merge(all, census, by.x = "city", by.y = "city_name")
+final_df <- merge(final_df, gun_provisions, by = "state")
+final_df <- 
 ```
 
+```{r save-all}
+saveRDS(final_df, "data/final.rds")
+```
+
+### OLD CODE 
 ```{r}

### 2020-10-17 submissionii.Rmd
merging dfs

@@
 |        As a basis for this study, I will be using the same dataset that VICE News used.
-This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from 2010 to 2016, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
+This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from January 2010 to September 2016, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
 Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), Other (O), and Unknown (U).
@@
 A 1 in the gun provision column represents a presence of the law, and a 0 indicates an absence.
-I also will add a column to this gun provision dataset indicating whether a state requires de-escalation training based off of information reported by APM Report in 2017, with a 1 meaning the state does require training and a 0 meaning the state does not.[4]
+I will also add a column to this gun provision dataset indicating whether a state requires de-escalation training based off of information reported by APM Report in 2017, with a 1 meaning the state does require training and a 0 meaning the state does not.
+There were 16 states that required de-escalation training as of November 2017.
+[4]
 
@@
 VICE News most likely provided only 2013 census data because it is the average year of all the years represented in the shootings dataset.
+Furthermore, using census data only from 2013 requires that we analyze this data under the assumption that there is no drastic difference in population demographics between 2010 and 2016.
 The census data includes 7 variables: the police department and the city it's located in, as well as the city's Black, Asian, Hispanic, White, and overall total population in 2013.
@@
 Finally, I filtered out all rows that had only unknown values for the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether the shooting was fatal or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
+The cleaned dataset contains a total of 2,084 observations.
 
-In the VICE dataset, I will be filtering out the cases in which fatality is unknown, or the subject’s or officer’s race or gender is unknown from the final dataset.
-Additionally, there are 265 rows in which there were more than one victims of the shooting; in these cases, I will be separating each of these observations into multiple rows, with each row representing each individual victim and their associated information.
-I will carry out the same procedure in the cases in which there were more than one officers present at the shooting.
-Because there can be multiple victims and officers recorded for a shooting event, I will also be filtering out cases in which for a shooting event, the number of observations listed for victim or officer race differs from the number of observations listed for victim or officer gender, since I can not infer the missing values in either of these columns.
+|        For the gun provisions dataset, I created a new column that sums up the total number of laws listed in the dataset that each state had in 2013 (for the same reason that VICE News used 2013).
+Additionally, as I mentioned earlier, I appended a column to this dataset representing whether a state requires de-escalation training for police officers as of November 2017.
+2017 was used under the assumption that no major event occurred between September 2016 and November 2017 that caused a sudden increase in the number of states that require de-escalation training for police officers.
+Finally, for the census data, the only cleaning that needed to be done was extracting the state that each city was in from the department column so that the census could be easily joined with the cleaned shootings dataset by city name, as well as the gun provisions dataset by abbreviated state name.
 
-|        After cleaning the data, I was left with a total of 2,084 observations.
+|        The final dataset consists of
 

### 2020-10-17 police_eda.Rmd
individual

@@
          unarmed = as.numeric(weapon == "unarmed"),
-         unknown_weapon = as.numeric(weapon == "U")) %>% 
-  select(-c(weapon, SubjectRace, SubjectGender, OfficerRace, OfficerGender))
+         unknown_weapon = as.numeric(weapon == "U"),
+         fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
+  select(-c(weapon, SubjectGender, OfficerRace, OfficerGender))
 
@@
 all <- all[-which(all$ob == 0 & all$ow == 0 & all$ol == 0 & all$oa == 0),]
+all <- all[-which(str_count(all$Fatal, "F") == 0 & str_count(all$Fatal, "N") == 0),]
+all <- all %>% 
+  select(-c(Fatal))
 ```
 
-```{r}
+```{r merge-data}
 final_df <- merge(all, census, by.x = "city", by.y = "city_name")
 final_df <- merge(final_df, gun_provisions, by = "state")
-final_df <- 
+final_df <- final_df %>% 
+  select(-c(department, year))
 ```
 
-```{r save-all}
+```{r save-final-df}
 saveRDS(final_df, "data/final.rds")

### 2020-10-17 submissionii.Rmd
individual

@@
 author: "Connie Wu"
-date: "10/12/2020"
+date: "10/17/2020"
 output:
@@
 ```{r load-data}
-# Census Predictors
-census <- read.csv("data/2013census.csv")[,-c(8,9)]
-states <- lapply(census$department, function(x) str_extract_all(x,"\\([^()]+\\)")[[1]])
-states <- substring(states, 2, nchar(states)-1)
-census$state <- states
-
-# Gun Law Predictors
-gun_provisions <- read.csv("data/gun_provisions.csv") %>% 
-  filter(year == 2013)
-gun_provisions <- gun_provisions %>% 
-  mutate(total_pr = rowSums(gun_provisions[,3:136]),
-         deesc_train = ifelse(state %in% c("California", "New Mexico", "Texas", "Oklahoma", "South Dakota", "Illinois", "Indiana", "Tennessee", "Washington", "Missouri", "Georgia", "Ohio", "Maryland", "Massachusetts", "Connecticut", "Maine"), 1, 0)) %>%
-  select(state, total_pr, deesc_train)
-gun_provisions$state <- state.abb[match(gun_provisions$state,state.name)]
-
-# Shootings
-all <- read.csv("data/allshootingsdata.csv")
-```
-
-```{r clean-all}
-# Recode officer races
-all$OfficerRace <- gsub("H/L", "L", all$OfficerRace)
-all$OfficerRace <- gsub("H", "L", all$OfficerRace)
-all$OfficerRace <- gsub("Unknown|m/m", "U", all$OfficerRace)
-all$OfficerRace <- gsub("WHITE|WLITE", "W", all$OfficerRace)
-all$OfficerRace <- gsub("BLACK", "B", all$OfficerRace)
-all$OfficerRace <- gsub("ASIAN", "A", all$OfficerRace)
-all$OfficerRace <- gsub("Other|AI/AN|NA/W|Multi-Racial", "O", all$OfficerRace)
-all$OfficerRace <- gsub("A/PI ", "A;", all$OfficerRace)
-all$OfficerRace <- gsub("A/PI", "A", all$OfficerRace)
-all$OfficerRace <- gsub(",", ";", all$OfficerRace)
-all$OfficerRace <- gsub(" ", "", all$OfficerRace)
-all$OfficerRace <- gsub("A/W|W/L|W/A", "O", all$OfficerRace)
-all$OfficerRace[all$OfficerRace == ""] <- "U"
-
-# Recode officer genders
-all$OfficerGender <- gsub("MALE", "M", toupper(all$OfficerGender))
-all$OfficerGender <- gsub("FEMALE|FEM", "F", toupper(all$OfficerGender))
-all$OfficerGender <- gsub("UNKNOWN|FALSE", "U", toupper(all$OfficerGender))
-all$OfficerGender <- gsub(",|:", ";", toupper(all$OfficerGender))
-all$OfficerGender <- gsub(" |/", "", toupper(all$OfficerGender))
-
-all <- all %>% 
-  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "tr", "AvgAge", "VA", "F", "nshots", "sit")) %>% 
-  mutate(b = as.numeric(b > 0),
-         w = as.numeric(w > 0),
-         l = as.numeric(l > 0),
-         a = as.numeric(a > 0),
-         ob = as.numeric(str_count(all$OfficerRace, "B") > 0),
-         ow = as.numeric(str_count(all$OfficerRace, "W") > 0),
-         ol = as.numeric(str_count(all$OfficerRace, "L") > 0),
-         oa = as.numeric(str_count(all$OfficerRace, "A") > 0),
-         other = as.numeric(other), 
-         replica = as.numeric(replica), 
-         knife = as.numeric(knife), 
-         gun = as.numeric(weapon == "gun"), 
-         unarmed = as.numeric(weapon == "unarmed"),
-         unknown_weapon = as.numeric(weapon == "U"),
-         id = paste0("C", row_number())) %>% 
-  select(-c(weapon))
-
-all <- rename(all, sb = b, sl = l, sw = w, sa = a)
-
-all <- all %>% 
-  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")
-
-officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
-officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
-out <- which((officer_gender == officer_race) == FALSE)
-all <- all[-out,]
-
-all <- all %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ",")
-
-officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
-officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
-out2 <- which((officer_gender2 == officer_race2) == FALSE)
-all <- all[-out2,]
-
-all$OfficerRace <- trimws(all$OfficerRace)
-all$OfficerGender <- trimws(all$OfficerGender)
-all$Fatal <- trimws(all$Fatal)
-
-all <- all %>% 
-  filter(SubjectRace != "U") %>% 
-  filter(OfficerRace != "U") %>% 
-  filter(OfficerRace != "Unknown") %>% 
-  filter(OfficerRace != "NA") %>% 
-  filter(OfficerRace != "") %>% 
-  filter(OfficerRace != "I") %>% 
-  filter(OfficerGender != "U") %>% 
-  filter(OfficerGender != "N") %>% 
-  filter(SubjectGender != "U") %>% 
-  filter(SubjectGender != "N") %>% 
-  filter(Fatal != "U")
-
-all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
-all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
-all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"
-
-all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
-all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
-all$OfficerGender[all$OfficerGender == "Male"] <- "M"
+final_df <- readRDS("data/final.rds")
 ```
@@
 |        After this initial cleaning, I created new columns in the same dataset that represented whether any victims of race Black, White, Asian, or Other were present in the crime, respectively, and if any male or female victims were present in the crime, respectively.
-I also added new columns to represent the same information for officers (for each race and gender, whether each was present).
-I then made sure that each weapon type had its own column.
+I also added new columns to represent the same information for officers (for each race and gender, whether each was present), and a new column to represent whether any victim involved in the shooting was shot fatally.
+Additionally, I made sure that each weapon type had its own column.
 All of these added columns had values of either 0 or 1, with 0 representing an absence of the variable and 1 a presence.
-Finally, I filtered out all rows that had only unknown values for the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether the shooting was fatal or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
-The cleaned dataset contains a total of 2,084 observations.
+Finally, I filtered out 2,079 rows that had only unknown values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
 
@@
 
-|        The final dataset consists of
+|        After merging the three datasets, the final dataset consists of 2,038 observations and 34 columns.
 
@@
 
-```{r}
-single_victims <- all %>% 
+```{r num_rows, eval = F}
+single_victims <- final_df %>% 
   filter(NumberOfSubjects <= 1)
@@
 
-mult_victims <- all %>% 
+mult_victims <- final_df %>% 
   filter(NumberOfSubjects > 1)
 nrow(mult_victims)
+```
 
-unknown <- all %>% 
-  filter(str_detect(SubjectRace, "U"))
-```
-|       Upon first glance, there appears to be 1932 incidents in which there was only one victim, and 152 with multiple victims.
-This means that there is one incident in which the victim race was marked as "NA." Additionally,
-
-```{r eval = F}
-single_victims <- all %>% 
-  filter(!str_detect(SubjectRace, ";"))
-
-mult_victims <- all %>% 
-  filter(str_detect(SubjectRace, ";"))
-
+```{r, eval = F}
 ggplot(single_victims, aes(x = Fatal, y = count)) +
@@
 
-```{r fatal-eda2}
+|       Upon first glance, there appears to be 1,890 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,
+
+```{r fatal-eda2, eval = F}
 temp <- all %>% 
@@
 
-```{r race-eda}
+```{r race-eda, eval = F}
 temp <- all %>% 
@@
 
-```{r gender-eda}
+```{r gender-eda, eval = F}
 temp <- all %>% 
@@
 ### Model 
-The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
+|        The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
 The response I will be modeling will be whether the shooting was fatal or not.
@@
 
+- limitations: time series, also counting N;U... combos as N even though some are unknown
 ## References 

### 2020-10-18 police_eda.Rmd
individual

@@
 # Recode subject race
-all$SubjectRace[is.na(all$SubjectRace)] <- "U"
+all$SubjectRace[is.na(all$SubjectRace)] <- "A"
 
@@
 all <- all[-which(all$sb == 0 & all$sw == 0 & all$sl == 0 & all$sa == 0),]
+all <- all[-which(is.na(all$sb)),]
 all <- all[-which(all$ob == 0 & all$ow == 0 & all$ol == 0 & all$oa == 0),]

### 2020-10-18 submissionii.Rmd
individual

@@
 All of these added columns had values of either 0 or 1, with 0 representing an absence of the variable and 1 a presence.
-Finally, I filtered out 2,079 rows that had only unknown values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
+Finally, I filtered out 2,079 rows that had only unknown and NA values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
 
@@
 
-|        After merging the three datasets, the final dataset consists of 2,038 observations and 34 columns.
+|        After merging the three datasets, the final dataset consists of 2,031 observations.
 
@@
 ```{r num_rows, eval = F}
-single_victims <- final_df %>% 
+final_eda <- final_df %>%
+  mutate(
+      fatal = case_when(
+        fatal == 1 ~ "Fatal",
+        fatal == 0 ~ "Non-Fatal"
+    )
+  )
+
+single_victims <- final_eda %>% 
   filter(NumberOfSubjects <= 1)
@@
 
-mult_victims <- final_df %>% 
+mult_victims <- final_eda %>% 
   filter(NumberOfSubjects > 1)
@@
 ```{r, eval = F}
-ggplot(single_victims, aes(x = Fatal, y = count)) +
+tempsb <- single_victims %>% 
+  group_by(sb, fatal) %>% 
+  summarise(count = n())
+tempsw <- single_victims %>% 
+  group_by(sw, fatal) %>% 
+  summarise(count = n())
+tempsa <- single_victims %>% 
+  group_by(sa, fatal) %>% 
+  summarise(count = n())
+tempsl <- single_victims %>% 
+  group_by(sl, fatal) %>% 
+  summarise(count = n())
+
+## why are there no others?
+tempso <- single_victims %>% 
+  group_by(so, fatal) %>% 
+  summarise(count = n())
+
+ggplot(single_victims, aes(x = fatal, y = count)) +
   geom_bar(stat="identity") + 
@@
 
-|       Upon first glance, there appears to be 1,890 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,
+|       Upon first glance, there appears to be 1,883 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,
 

### 2020-10-18 police_eda.Rmd
fixed some stuff

@@
          ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
-         other = as.numeric(other), 
+         weapon_other = as.numeric(other), 
          replica = as.numeric(replica), 
@@
 all <- all[-which(all$omale == 0 & all$ofem == 0),]
-all <- all[-which(all$sb == 0 & all$sw == 0 & all$sl == 0 & all$sa == 0),]
+all <- all[-which(all$sb == 0 & all$sw == 0 & all$sl == 0 & all$sa == 0 & all$so == 0),]
 all <- all[-which(is.na(all$sb)),]
-all <- all[-which(all$ob == 0 & all$ow == 0 & all$ol == 0 & all$oa == 0),]
+all <- all[-which(all$ob == 0 & all$ow == 0 & all$ol == 0 & all$oa == 0 & all$oo == 0),]
 all <- all[-which(str_count(all$Fatal, "F") == 0 & str_count(all$Fatal, "N") == 0),]
@@
 
+```{r single_victims-race}
+tempsb <- single_victims %>% 
+  group_by(sb, fatal) %>% 
+  summarise(count = n())
+
+ggplot(tempsb, aes(x = sb, y = count, fill = fatal)) +
+  geom_bar(stat="identity") + 
+  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings")
+
+tempsw <- single_victims %>% 
+  group_by(sw, fatal) %>% 
+  summarise(count = n())
+tempsa <- single_victims %>% 
+  group_by(sa, fatal) %>% 
+  summarise(count = n())
+tempsl <- single_victims %>% 
+  group_by(sl, fatal) %>% 
+  summarise(count = n())
+tempso <- single_victims %>% 
+  group_by(so, fatal) %>% 
+  summarise(count = n())
+
+```
+
 ```{r}

### 2020-10-18 submissionii.Rmd
fixed some stuff

@@
         fatal == 1 ~ "Fatal",
-        fatal == 0 ~ "Non-Fatal"
+        fatal == 0 ~ "Non-Fatal"),
+      sb = case_when(
+        sb == 1 ~ "Victim Type Present",
+        sb == 0 ~ "Victim Type Not Present"
+      ),
+      sw = case_when(
+        sw == 1 ~ "Victim Type Present",
+        sw == 0 ~ "Victim Type Not Present"
+      ),
+      sl = case_when(
+        sl == 1 ~ "Victim Type Present",
+        sl == 0 ~ "Victim Type Not Present"
+      ),
+      sa = case_when(
+        sa == 1 ~ "Victim Type Present",
+        sa == 0 ~ "Victim Type Not Present"
+      ),
+      so = case_when(
+        so == 1 ~ "Victim Type Present",
+        so == 0 ~ "Victim Type Not Present"
+      ),
+      ob = case_when(
+        ob == 1 ~ "Officer Type Present",
+        ob == 0 ~ "Officer Type Not Present"
+      ),
+      ow = case_when(
+        ow == 1 ~ "Officer Type Present",
+        ow == 0 ~ "Officer Type Not Present"
+      ),
+      ol = case_when(
+        ol == 1 ~ "Officer Type Present",
+        ol == 0 ~ "Officer Type Not Present"
+      ),
+      oa = case_when(
+        oa == 1 ~ "Officer Type Present",
+        oa == 0 ~ "Officer Type Not Present"
+      ),
+      oo = case_when(
+        oo == 1 ~ "Officer Type Present",
+        oo == 0 ~ "Officer Type Not Present"
       )
@@
   summarise(count = n())
-
-## why are there no others?
 tempso <- single_victims %>% 
@@
 
-|       Upon first glance, there appears to be 1,883 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,
+|       Upon first glance, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,
 

### 2020-10-18 police_eda.Rmd
adding eda

@@
          ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
-         weapon_other = as.numeric(other), 
+         other = as.numeric(other), 
          replica = as.numeric(replica), 
@@
          fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
-  select(-c(weapon, SubjectGender, OfficerRace, OfficerGender))
+  select(-c(weapon, SubjectGender, OfficerGender))
 
-all <- rename(all, sb = b, sw = w, sl = l, sa = a)
+all <- rename(all, sb = b, sw = w, sl = l, sa = a, weapon_other = other)
 ```
@@
 
-### OLD CODE 
-```{r}
-all <- all %>% 
-  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")
+## Exploratory Data Analysis
 
-officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
-officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
-out <- which((officer_gender == officer_race) == FALSE)
-all <- all[-out,]
+```{r rename 0 and 1 values}
+final_eda <- final_df %>%
+  mutate(
+      fatal = case_when(
+        fatal == 1 ~ "Fatal",
+        fatal == 0 ~ "Non-Fatal"),
+      sb = case_when(
+        sb == 1 ~ "Black Victim Present",
+        sb == 0 ~ "Black Victim Not Present"
+      ),
+      sw = case_when(
+        sw == 1 ~ "White Victim Present",
+        sw == 0 ~ "White Victim Not Present"
+      ),
+      sl = case_when(
+        sl == 1 ~ "Latino Victim Present",
+        sl == 0 ~ "Latino Victim Not Present"
+      ),
+      sa = case_when(
+        sa == 1 ~ "Asian Victim Present",
+        sa == 0 ~ "Asian Victim Not Present"
+      ),
+      so = case_when(
+        so == 1 ~ "Other Victim Present",
+        so == 0 ~ "Other Victim Not Present"
+      ),
+      smale = case_when(
+        smale == 1 ~ "Male Victim Present",
+        smale == 0 ~ "Male Victim Not Present"
+      ),
+      sfem = case_when(
+        sfem == 1 ~ "Female Victim Present",
+        sfem == 0 ~ "Female Victim Not Present"
+      ),
+      ob = case_when(
+        ob == 1 ~ "Black Officer Present",
+        ob == 0 ~ "Black Officer Not Present"
+      ),
+      ow = case_when(
+        ow == 1 ~ "White Officer Present",
+        ow == 0 ~ "White Officer Not Present"
+      ),
+      ol = case_when(
+        ol == 1 ~ "Latino Officer Present",
+        ol == 0 ~ "Latino Officer Not Present"
+      ),
+      oa = case_when(
+        oa == 1 ~ "Asian Officer Present",
+        oa == 0 ~ "Asian Officer Not Present"
+      ),
+      oo = case_when(
+        oo == 1 ~ "Other Officer Present",
+        oo == 0 ~ "Other Officer Not Present"
+      ),
+      omale = case_when(
+        omale == 1 ~ "Male Officer Present",
+        omale == 0 ~ "Male Officer Not Present"
+      ),
+      weapon_other = case_when(
+        weapon_other == 1 ~ "Other Weapon Present",
+        weapon_other == 0 ~ "Other Weapon Not Present"
+      ), 
+      replica = case_when(
+        replica == 1 ~ "Replica Present",
+        replica == 0 ~ "Replica Not Present"
+      ),
+      knife = case_when(
+        knife == 1 ~ "Knife Present",
+        knife == 0 ~ "Knife Not Present"
+      ),
+      gun = case_when(
+        gun == 1 ~ "Gun Present",
+        gun == 0 ~ "Gun Not Present"
+      ),
+      unarmed = case_when(
+        unarmed == 1 ~ "Victim Unarmed",
+        unarmed == 0 ~ "Victim Armed"
+      ),
+      unknown_weapon = case_when(
+        unknown_weapon == 1 ~ "Weapon Unknown",
+        unknown_weapon == 0 ~ "Weapon Known"
+      )
+  )
 
-all <- all %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ",")
-
-officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
-officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
-out2 <- which((officer_gender2 == officer_race2) == FALSE)
-all <- all[-out2,]
-
-all$OfficerRace <- trimws(all$OfficerRace)
-all$OfficerGender <- trimws(all$OfficerGender)
-all$Fatal <- trimws(all$Fatal)
-
-all <- all %>% 
-  filter(SubjectRace != "U") %>% 
-  filter(OfficerRace != "U") %>% 
-  filter(OfficerRace != "Unknown") %>% 
-  filter(OfficerRace != "NA") %>% 
-  filter(OfficerRace != "") %>% 
-  filter(OfficerRace != "I") %>% 
-  filter(OfficerGender != "U") %>% 
-  filter(OfficerGender != "N") %>% 
-  filter(SubjectGender != "U") %>% 
-  filter(SubjectGender != "N") %>% 
-  filter(Fatal != "U")
-
-all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
-all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
-all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"
-
-all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
-all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
-all$OfficerGender[all$OfficerGender == "Male"] <- "M"
-
+saveRDS(final_eda, 'data/final_eda.rds')
 ```
 
-```{r split_data}
-single_victims <- all %>% 
-  filter(!str_detect(SubjectRace, ";"))
+```{r race shootings vs population demographics}
+most_pop <- tail(names(sort(table(final_eda$city))), 6)
 
-mult_victims <- all %>% 
-  filter(str_detect(SubjectRace, ";"))
+for (c in most_pop) {
+  temp <- final_eda %>% 
+    filter(city == c)
+
+  vic <- temp %>% 
+  summarise(total_b = sum(str_count(SubjectRace, "B")), 
+            total_w = sum(str_count(SubjectRace, "W")),
+            total_l = sum(str_count(SubjectRace, "L")),
+            total_a = sum(str_count(SubjectRace, "A")),
+            total_o = sum(str_count(SubjectRace, "O")))
+
+  vic <- vic %>%
+    t() %>%
+    as.data.frame(stringsAsFactors = F) %>%
+    rownames_to_column("value") %>%
+    `rownames<-`(NULL)
+
+  demo <-
+
+}
+
+vic <- final_eda %>% 
+  summarise(total_b = sum(str_count(SubjectRace, "B")), 
+            total_w = sum(str_count(SubjectRace, "W")),
+            total_l = sum(str_count(SubjectRace, "L")),
+            total_a = sum(str_count(SubjectRace, "A")),
+            total_o = sum(str_count(SubjectRace, "O")))
+vic <- vic %>%
+  t() %>%
+  as.data.frame(stringsAsFactors = F) %>%
+  rownames_to_column("value") %>%
+  `rownames<-`(NULL)
+
+pop <- final_eda %>% 
+  summarise(total_b = sum())
 ```
 
-## Exploratory Data Analysis
+```{r split into single vs mult}
+single_victims <- final_eda %>% 
+  filter(NumberOfSubjects <= 1)
+nrow(single_victims)
 
-```{r single_victims-race}
+mult_victims <- final_eda %>% 
+  filter(NumberOfSubjects > 1)
+nrow(mult_victims)
+```
+
+```{r single victims race counts and props}
 tempsb <- single_victims %>% 
@@
   summarise(count = n())
-
-ggplot(tempsb, aes(x = sb, y = count, fill = fatal)) +
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings")
+tempsb <- tempsb[3:4,]
+tempsb <- rename(tempsb, type = sb)
 
@@
   summarise(count = n())
+tempsw <- tempsw[3:4,]
+tempsw <- rename(tempsw, type = sw)
+
 tempsa <- single_victims %>% 
@@
   summarise(count = n())
+tempsa <- tempsa[3:4,]
+tempsa <- rename(tempsa, type = sa)
+
 tempsl <- single_victims %>% 
@@
   summarise(count = n())
+tempsl <- tempsl[3:4,]
+tempsl <- rename(tempsl, type = sl)
+
 tempso <- single_victims %>% 
@@
   summarise(count = n())
+tempso <- tempso[3:4,]
+tempso <- rename(tempso, type = so)
 
+temp <- rbind(tempsb, tempsw, tempsl, tempsa, tempso) %>% 
+  group_by(type) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  mutate(prop = count / total)
+
+race_count <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
+  geom_bar(position = "dodge", stat="identity") + 
+  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
+  theme(legend.text = element_text(size = 8),
+        legend.key.size = unit(0.4, "cm"),
+        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        plot.title = element_text(size=10, hjust = 0.5)) + 
+  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))
+
+race_prop <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
+  geom_bar(position = "stack", stat="identity") + 
+  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
+  theme(legend.text = element_text(size = 7),
+        legend.key.size = unit(0.4, "cm"),
+        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        plot.title = element_text(size=10, hjust = 0.5)) + 
+  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))
+
+ggsave(filename = "figures/eda1.png", race_count)
+ggsave(filename = "figures/eda2.png", race_prop)
+```
+
+```{r multiple victims race counts and props}
+tempsb <- mult_victims %>% 
+  group_by(sb, fatal) %>% 
+  summarise(count = n())
+tempsb <- tempsb[3:4,]
+tempsb <- rename(tempsb, type = sb)
+
+tempsw <- mult_victims %>% 
+  group_by(sw, fatal) %>% 
+  summarise(count = n())
+tempsw <- tempsw[3:4,]
+tempsw <- rename(tempsw, type = sw)
+
+tempsa <- mult_victims %>% 
+  group_by(sa, fatal) %>% 
+  summarise(count = n())
+tempsa <- tempsa[3,]
+tempsa[2,] <- list("Asian Victim Present", "Fatal", 0)
+tempsa <- rename(tempsa, type = sa)
+
+tempsl <- mult_victims %>% 
+  group_by(sl, fatal) %>% 
+  summarise(count = n())
+tempsl <- tempsl[3:4,]
+tempsl <- rename(tempsl, type = sl)
+
+temp <- rbind(tempsb, tempsw, tempsl, tempsa) %>% 
+  group_by(type) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  mutate(prop = count / total)
+
+race_count1 <- ggplot(temp, aes(x = type, y = count, fill = fatal)) +
+  geom_bar(position = "dodge", stat="identity") + 
+  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
+  theme(legend.text = element_text(size = 8),
+        legend.key.size = unit(0.4, "cm"),
+        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        plot.title = element_text(size=10, hjust = 0.5)) + 
+  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))
+
+race_prop1 <- ggplot(temp, aes(x = type, y = prop, fill = fatal)) +
+  geom_bar(position = "stack", stat="identity") + 
+  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
+  theme(legend.text = element_text(size = 7),
+        legend.key.size = unit(0.4, "cm"),
+        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        plot.title = element_text(size=10, hjust = 0.5)) + 
+  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))
+
+ggsave(filename = "figures/eda3.png", race_count1)
+ggsave(filename = "figures/eda4.png", race_prop1)
 ```
@@
 ```
+
+### OLD CODE 
+```{r}
+all <- all %>% 
+  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")
+
+officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
+officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
+out <- which((officer_gender == officer_race) == FALSE)
+all <- all[-out,]
+
+all <- all %>% 
+  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
+  separate_rows(OfficerRace, OfficerGender, sep = ",")
+
+officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
+officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
+out2 <- which((officer_gender2 == officer_race2) == FALSE)
+all <- all[-out2,]
+
+all$OfficerRace <- trimws(all$OfficerRace)
+all$OfficerGender <- trimws(all$OfficerGender)
+all$Fatal <- trimws(all$Fatal)
+
+all <- all %>% 
+  filter(SubjectRace != "U") %>% 
+  filter(OfficerRace != "U") %>% 
+  filter(OfficerRace != "Unknown") %>% 
+  filter(OfficerRace != "NA") %>% 
+  filter(OfficerRace != "") %>% 
+  filter(OfficerRace != "I") %>% 
+  filter(OfficerGender != "U") %>% 
+  filter(OfficerGender != "N") %>% 
+  filter(SubjectGender != "U") %>% 
+  filter(SubjectGender != "N") %>% 
+  filter(Fatal != "U")
+
+all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
+all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
+all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"
+
+all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
+all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
+all$OfficerGender[all$OfficerGender == "Male"] <- "M"
+```

### 2020-10-18 submissionii.Rmd
adding eda

--- 2020-10-18 submissionii.Rmd
@@
 
+|       There appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
+
 ```{r num_rows, eval = F}
-final_eda <- final_df %>%
-  mutate(
-      fatal = case_when(
-        fatal == 1 ~ "Fatal",
-        fatal == 0 ~ "Non-Fatal"
-    )
-  )
+final_eda <- readRDS('data/final_eda.rds')
 
@@
 
-```{r, eval = F}
-tempsb <- single_victims %>% 
-  group_by(sb, fatal) %>% 
-  summarise(count = n())
-tempsw <- single_victims %>% 
-  group_by(sw, fatal) %>% 
-  summarise(count = n())
-tempsa <- single_victims %>% 
-  group_by(sa, fatal) %>% 
-  summarise(count = n())
-tempsl <- single_victims %>% 
-  group_by(sl, fatal) %>% 
-  summarise(count = n())
-
-## why are there no others?
-tempso <- single_victims %>% 
-  group_by(so, fatal) %>% 
-  summarise(count = n())
-
-ggplot(single_victims, aes(x = fatal, y = count)) +
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-  group_by(Fatal, race) %>% 
-  summarise(count = n())
+```{r, figures-race, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
+par(mar = c(2, 2, .1, .1))
+knitr::include_graphics("figures/eda1.png")
+knitr::include_graphics("figures/eda2.png")
 ```
 
-|       Upon first glance, there appears to be 1,883 incidents in the final dataset in which there was only one victim, and 148 with multiple victims. Among these,
+|       Among all shootings with only one victim, it appears that the victim is more often Black than some other race.
+However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally.
+Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small.
+This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.
 
@@
 - limitations: time series, also counting N;U... combos as N even though some are unknown
+
 ## References 
@@
 4) http://govred.com/blog/deescalation-training-state-requirements/
+
+## Appendix

### 2020-10-18 police_eda.Rmd
adding more eda

@@
 library(tidyverse)
+library(reshape2)
 ```
@@
 ```{r race shootings vs population demographics}
-most_pop <- tail(names(sort(table(final_eda$city))), 6)
+most_pop <- head(names(sort(table(final_eda$city), decreasing = TRUE)), 4)
+i <- 1
 
@@
   vic <- temp %>% 
-  summarise(total_b = sum(str_count(SubjectRace, "B")), 
-            total_w = sum(str_count(SubjectRace, "W")),
-            total_l = sum(str_count(SubjectRace, "L")),
-            total_a = sum(str_count(SubjectRace, "A")),
-            total_o = sum(str_count(SubjectRace, "O")))
+  summarise(Black = sum(str_count(SubjectRace, "B")), 
+            White = sum(str_count(SubjectRace, "W")),
+            Hispanic = sum(str_count(SubjectRace, "L")),
+            Asian = sum(str_count(SubjectRace, "A")),
+            city_pop_2013 = sum(str_count(SubjectRace, "B")) + sum(str_count(SubjectRace, "W")) + 
+              sum(str_count(SubjectRace, "L")) + sum(str_count(SubjectRace, "A")) + 
+              sum(str_count(SubjectRace, "O"))
+  ) %>% 
+    mutate(Black = Black/city_pop_2013, 
+           White = White/city_pop_2013,
+           Hispanic = Hispanic/city_pop_2013,
+           Asian = Asian/city_pop_2013)
 
@@
     as.data.frame(stringsAsFactors = F) %>%
-    rownames_to_column("value") %>%
+    rownames_to_column("race") %>%
     `rownames<-`(NULL)
 
-  demo <-
+  demo <- temp %>% 
+    select(black, white, hispanic, asian, city_pop_2013)  %>% 
+    mutate(black = black/city_pop_2013, 
+           white = white/city_pop_2013,
+           hispanic = hispanic/city_pop_2013,
+           asian = asian/city_pop_2013)
+  demo <- rename(demo, Black = black, White = white, Hispanic = hispanic, Asian = asian)
+  demo <- demo[1,]
 
-}
-
-vic <- final_eda %>% 
-  summarise(total_b = sum(str_count(SubjectRace, "B")), 
-            total_w = sum(str_count(SubjectRace, "W")),
-            total_l = sum(str_count(SubjectRace, "L")),
-            total_a = sum(str_count(SubjectRace, "A")),
-            total_o = sum(str_count(SubjectRace, "O")))
-vic <- vic %>%
+  demo <- demo %>%
     t() %>%
     as.data.frame(stringsAsFactors = F) %>%
-  rownames_to_column("value") %>%
+    rownames_to_column("race") %>%
     `rownames<-`(NULL)
+  demo <- rename(demo, V1 = 2)
 
-pop <- final_eda %>% 
-  summarise(total_b = sum())
+  temp2 <- merge(vic, demo, by = "race") %>% 
+    filter(race != "city_pop_2013")
+  temp2 <- rename(temp2, "Shootings" = V1.x, "City-Wide" = V1.y)
+  temp2 <- melt(temp2)
+
+  plot <- ggplot(temp2, aes(x = race, y = value, fill = variable)) +
+    geom_bar(position = "dodge", stat="identity") + 
+    labs(x = "Victim Race", y = "Proportion", title = paste0("Race Demographics of Shooting Victims vs. ", c, " Population in 2013")) +
+    theme(legend.text = element_text(size = 8),
+          legend.key.size = unit(0.4, "cm"),
+          plot.title = element_text(size=10, hjust = 0.5)) + 
+    scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))
+
+  ggsave(filename = paste0("figures/cityeda", i, ".png"), plot)
+
+  i <- i + 1
+}
 ```
@@
 
-ggsave(filename = "figures/eda1.png", race_count)
-ggsave(filename = "figures/eda2.png", race_prop)
+ggsave(filename = "figures/singleeda1.png", race_count)
+ggsave(filename = "figures/singleeda2.png", race_prop)
 ```
@@
 
-ggsave(filename = "figures/eda3.png", race_count1)
-ggsave(filename = "figures/eda4.png", race_prop1)
+ggsave(filename = "figures/multeda1.png", race_count1)
+ggsave(filename = "figures/multeda2.png", race_prop1)
 ```

### 2020-10-18 submissionii.Rmd
adding more eda

@@
 
-|       There appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
+|
+
+```{r, figures-city, fig.show="hold", out.width="22%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
+par(mar = c(3, 3, .1, .1))
+knitr::include_graphics("figures/cityeda1.png")
+knitr::include_graphics("figures/cityeda2.png")
+knitr::include_graphics("figures/cityeda3.png")
+knitr::include_graphics("figures/cityeda4.png")
+```
+
+|       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
 
@@
 
-```{r, figures-race, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
+```{r, figures-single, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
 par(mar = c(2, 2, .1, .1))
-knitr::include_graphics("figures/eda1.png")
-knitr::include_graphics("figures/eda2.png")
+knitr::include_graphics("figures/singleeda1.png")
+knitr::include_graphics("figures/singleeda2.png")
 ```

### 2020-10-18 police_eda.Rmd
more eda

@@
 
+  if (c == "LosAngeles") {
+    c <- "Los Angeles"
+  }
+
   plot <- ggplot(temp2, aes(x = race, y = value, fill = variable)) +
@@
           legend.key.size = unit(0.4, "cm"),
-          plot.title = element_text(size=10, hjust = 0.5)) + 
+          plot.title = element_text(size=13, hjust = 0.5)) + 
     scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))

### 2020-10-18 submissionii.Rmd
more eda

--- 2020-10-18 submissionii.Rmd
@@
 
-|       There appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
+|       Plotted below are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs.
+the race proportions of those cities' total populations in 2013.
+Because the bars representing shootings in the Black race category are taller than the bars representing city-wide population in all 4 plots, it is evident Black people are disproportionately the victims of police shootings compared to city race proportions, especially in Chicago and Houston.
+
+```{r, figures-city, fig.show="hold", out.width="30%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
+knitr::include_graphics("figures/cityeda1.png")
+knitr::include_graphics("figures/cityeda2.png")
+knitr::include_graphics("figures/cityeda3.png")
+knitr::include_graphics("figures/cityeda4.png")
+```
 
@@
 
-```{r, figures-race, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
+```{r, figures-single, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
 par(mar = c(2, 2, .1, .1))
-knitr::include_graphics("figures/eda1.png")
-knitr::include_graphics("figures/eda2.png")
+knitr::include_graphics("figures/singleeda1.png")
+knitr::include_graphics("figures/singleeda2.png")
 ```
 
-|       Among all shootings with only one victim, it appears that the victim is more often Black than some other race.
+|       There appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
+Among all shootings with only one victim, it appears that the victim is more often Black than some other race.
 However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally.

### 2020-10-18 submissionii.Rmd
eda

@@
 the race proportions of those cities' total populations in 2013.
-Because the bars representing shootings in the Black race category are taller than the bars representing city-wide population in all 4 plots, it is evident Black people are disproportionately the victims of police shootings compared to city race proportions, especially in Chicago and Houston.
+Because the bars representing shootings in the Black race category are taller than the bars representing city-wide population in all 4 plots, it is evident Black people are disproportionately the victims of police shootings compared to city race proportions.
+This can especially be seen in Houston, where Black people make up less than 25% of the city population, yet more than 50% of the victims of police shootings in Houston.
 
@@
 
-|       There appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
+|       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
 Among all shootings with only one victim, it appears that the victim is more often Black than some other race.
@@
 This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.
-
-```{r fatal-eda2, eval = F}
-temp <- all %>% 
-  group_by(Fatal) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=Fatal, y=count)) + 
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-```
-
-|        In this dataset, it is important to keep in mind that there is a higher number of non-fatal shooting events than fatal shooting events.
-
-```{r race-eda, eval = F}
-temp <- all %>% 
-  group_by(Fatal, SubjectRace) %>% 
-  summarise(race_count = n())
-
-ggplot(temp, aes(x=SubjectRace, y=race_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in Police Shootings")
-
-temp2 <- all %>% 
-  group_by(Fatal, OfficerRace) %>% 
-  summarise(race_count = n())
-
-ggplot(temp2, aes(x=OfficerRace, y=race_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-|        In the plots above, we can see the counts for both victim and officer race in the final dataset split by fatal versus non-fatal shootings.
-In this dataset, A stands for Asian, B stands for Black, H stands for Hispanic, L stands for Latino, O stands for other, and W stands for White.
-Already, we can see that in both fatal and non-fatal cases, Black is the race with the highest count for the victim category, whereas White is the race with the highest count for the officer category.
-It is also interesting to note that the number of Black victims is higher in non-fatal cases than in fatal cases, while the number of white victims is higher in fatal cases than in non-fatal cases.
-
-```{r gender-eda, eval = F}
-temp <- all %>% 
-  group_by(Fatal, SubjectGender) %>% 
-  summarise(gender_count = n())
-
-ggplot(temp, aes(x=SubjectGender, y=gender_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in Police Shootings")
-
-temp2 <- all %>% 
-  group_by(Fatal, OfficerGender) %>% 
-  summarise(gender_count = n())
-
-ggplot(temp2, aes(x=OfficerGender, y=gender_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-|        Additionally, it is evident from the graphs above that male is the gender with the highest count in the cases of victims and officers for both fatal and non-fatal shootings.
-
-```{r weapon-eda, eval=F}
-temp <- all %>% 
-  group_by(Fatal, weapon) %>% 
-  summarise(count = n())
-ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings")
-```
-
-|        Finally, we can see that in cases of police shootings, victims have guns the most often in fatal cases and unknown weapons in non-fatal cases, whereas they have gun replicas the least often in both cases.
 

### 2020-10-18 police_eda.Rmd
more eda

@@
   select(-c(department, year))
+final_df$NumberOfSubjects[which(final_df$NumberOfSubjects == 0)] <- 1
 ```
@@
 
-```{r}
-temp <- single_victims %>% 
-  group_by(Fatal, SubjectRace) %>% 
-  summarise(count = n())
-
-ggplot(single_victims, aes(x = Fatal, y = count)) +
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
-  group_by(Fatal, race) %>% 
-  summarise(count = n())
+```{r fatally shot prop for all shootings of varying numbers of victims}
+temp <- final_eda %>% 
+  group_by(NumberOfSubjects) %>% 
+  summarise(prop = sum(fatal == "Fatal")/n(),
+            total = n())
+temp
 ```
 
-```{r fatal-eda}
-temp <- all %>% 
-  group_by(Fatal) %>% 
-  summarise(count = n())
+```{r black to white ratio}
+temp <- final_eda %>% 
+  group_by(city, fatal) %>% 
+  summarise(btw = black/white) %>% 
+  unique
 
-ggplot(temp, aes(x=Fatal, y=count)) + 
-  geom_bar(stat="identity") + 
-  labs(x = "Fatal", y = "Count", title = "Counts for Fatal vs. Non-Fatal Police Shootings") 
+plot <- ggplot(temp, aes(x = btw, y = fatal, fill = fatal)) + 
+    geom_boxplot()  + 
+  labs(x = "Black to White Ratio", y = "Fatal vs. Non-Fatal Shooting", title = "City Black to White Ratio Distributions for Fatal vs. Non-Fatal Police Shootings") +
+  theme(legend.text = element_text(size = 8),
+        legend.key.size = unit(0.4, "cm"),
+        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        plot.title = element_text(size=10, hjust = 0.5))
+
+ggsave(filename = "figures/btweda.png", plot)
 ```
-
-```{r race-eda}
-temp <- all %>% 
-  group_by(Fatal, SubjectRace) %>% 
-  summarise(race_count = n())
-
-ggplot(temp, aes(x=SubjectRace, y=race_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Victim Race by Fatal vs. Non-Fatal in Police Shootings")
-
-temp2 <- all %>% 
-  group_by(Fatal, OfficerRace) %>% 
-  summarise(race_count = n())
-
-ggplot(temp2, aes(x=OfficerRace, y=race_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Race", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-```{r gender-eda}
-temp <- all %>% 
-  group_by(Fatal, SubjectGender) %>% 
-  summarise(gender_count = n())
-
-ggplot(temp, aes(x=SubjectGender, y=gender_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Victim Gender by Fatal vs. Non-Fatal in Police Shootings")
-
-temp2 <- all %>% 
-  group_by(Fatal, OfficerGender) %>% 
-  summarise(gender_count = n())
-
-ggplot(temp2, aes(x=OfficerGender, y=gender_count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Gender", y = "Count", title = "Counts for Officer Race by Fatal vs. Non-Fatal in Police Shootings") 
-```
-
-```{r weapon-eda}
-temp <- all %>% 
-  group_by(Fatal, weapon) %>% 
-  summarise(count = n())
-
-ggplot(temp, aes(x=weapon, y=count, fill = Fatal)) + 
-  geom_bar(position="dodge", stat="identity") + 
-  labs(x = "Weapon Type", y = "Count", title = "Counts of Each Weapon Type on Victims by Fatal vs. Non-Fatal in Police Shootings") 
-```
+Means are not that different, will probably just use percent black and percent white then.
 

### 2020-10-18 submissionii.Rmd
more eda

@@
 [3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
-
 |        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders, whether they are carrying a weapon, and the total number of victims in the crime, play in fatal versus non-fatal police shootings.
@@
 I will also add a predictor indicating whether the state in which the homicide occurred requires de-escalation training for police officers.
-Finally, I will explore how the ratio of Black to White population in each location relates to police use of deadly force.
+Finally, I will explore how the race demographics of each location relates to police use of deadly force.
 This will allow me to better understand how racism has manifested itself in America’s police system and determine whether current attempts to prevent police use of lethal force are effective or not.
@@
 Each of these scenarios was still represented within one row of the dataset, but semi-colons were present in the respective victim and officer columns, separating each individual's information from each other.
-
 |       Additionally, to assess the effect of different types of legislation on fatal vs.
@@
 [4]
-
 |       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets.
@@
 For example, "A/PI" values were taken to represent Asian/Pacific Islander and thus were replaced with "A", and "A/W" or other values with a "/" in them were typically taken to represent multi-racial individuals and as a result were replaced with "O." Similar procedures were carried out in the gender column for officers for values that were not "M", "F", or "U".
-
 |        After this initial cleaning, I created new columns in the same dataset that represented whether any victims of race Black, White, Asian, or Other were present in the crime, respectively, and if any male or female victims were present in the crime, respectively.
@@
 Finally, I filtered out 2,079 rows that had only unknown and NA values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
-
 |        For the gun provisions dataset, I created a new column that sums up the total number of laws listed in the dataset that each state had in 2013 (for the same reason that VICE News used 2013).
@@
 Finally, for the census data, the only cleaning that needed to be done was extracting the state that each city was in from the department column so that the census could be easily joined with the cleaned shootings dataset by city name, as well as the gun provisions dataset by abbreviated state name.
-
 |        After merging the three datasets, the final dataset consists of 2,031 observations.
@@
 
-|       Plotted below are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs.
+|       Plotted below in \emph{Figure 1} are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs.
 the race proportions of those cities' total populations in 2013.
 Because the bars representing shootings in the Black race category are taller than the bars representing city-wide population in all 4 plots, it is evident Black people are disproportionately the victims of police shootings compared to city race proportions.
-This can especially be seen in Houston, where Black people make up less than 25% of the city population, yet more than 50% of the victims of police shootings in Houston.
-
-```{r, figures-city, fig.show="hold", out.width="30%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
+This can especially be seen in Houston, where Black people make up less than 25% of the city population yet more than 50% of the victim population of police shootings.
+```{r, figures-city, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
 knitr::include_graphics("figures/cityeda1.png")
 knitr::include_graphics("figures/cityeda2.png")
-knitr::include_graphics("figures/cityeda3.png")
-knitr::include_graphics("figures/cityeda4.png")
 ```
@@
 ```
-
 |       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
-Among all shootings with only one victim, it appears that the victim is more often Black than some other race.
+Among all shootings with only one victim, it appears in \emph{Figure 2} that the victim is more often Black than some other race.
 However, it seems that Black victims and most victims of the other race types are more likely to be shot non-fatally than fatally, whereas White victims seem more likely to be shot fatally.
@@
 This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.
+|       Similarly, Black victims make up the highest proportion of victims in shootings with multiple victims.
+However, in these shootings, all race types are more likely to be shot non-fatally than fatally, with Latinos having the highest proportion of being shot fatally out of all other race types.
+Additionally, it should be noted that there are no victims with race type "Other" in shootings where there are multiple victims.
+These graphs can be found as \emph{Figure 3} in the Appendix.
 
 ### Model
+
 |        The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
@@
 ## Appendix
+
+### Exploratory Data Analysis
+```{r, figures-mult, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Multiple Victim Shootings"}
+par(mar = c(2, 2, .1, .1))
+knitr::include_graphics("figures/multeda1.png")
+knitr::include_graphics("figures/multeda2.png")
+```
+
+```{r, figures-btw, fig.show="hold", out.width="50%", fig.cap="City Black to White Ratios Distributions for Fatal vs. Non-Fatal Police Shootings"}
+par(mar = c(2, 2, .1, .1))
+knitr::include_graphics("figures/btweda.png")
+```
+
+Means are not that different, will probably just use percent black and percent white then.

### 2020-10-19 police_eda.Rmd
eda

@@
          fatal = as.numeric(str_count(all$Fatal, "F") > 0)) %>% 
-  select(-c(weapon, SubjectGender, OfficerGender))
+  select(-c(weapon))
 
@@
   geom_bar(position = "dodge", stat="identity") + 
-  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
+  labs(x = "Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
   theme(legend.text = element_text(size = 8),
@@
   geom_bar(position = "stack", stat="identity") + 
-  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
+  labs(x = "Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Single-Victim Police Shootings Across Different Races") +
   theme(legend.text = element_text(size = 7),
@@
   geom_bar(position = "dodge", stat="identity") + 
-  labs(x = "Present Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
+  labs(x = "Victim Race", y = "Count", title = "Counts for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
   theme(legend.text = element_text(size = 8),
@@
   geom_bar(position = "stack", stat="identity") + 
-  labs(x = "Present Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
+  labs(x = "Victim Race", y = "Proportion", title = "Proportions for Fatal vs. Non-Fatal Multiple-Victim Police Shootings Across Different Races") +
   theme(legend.text = element_text(size = 7),
@@
 ```
+For sample size > 10, fatality seems to be lower for multiple victims rather than just 1.
 
 ```{r black to white ratio}
-temp <- final_eda %>% 
-  group_by(city, fatal) %>% 
-  summarise(btw = black/white) %>% 
+temp <- final_df %>% 
+  group_by(city) %>% 
+  summarise(btw = black/white, propfatal = sum(fatal)/n()) %>% 
   unique
 
-plot <- ggplot(temp, aes(x = btw, y = fatal, fill = fatal)) + 
-    geom_boxplot()  + 
-  labs(x = "Black to White Ratio", y = "Fatal vs. Non-Fatal Shooting", title = "City Black to White Ratio Distributions for Fatal vs. Non-Fatal Police Shootings") +
+plot <- ggplot(temp, aes(x = btw, y = propfatal)) + 
+    geom_point()  + 
+  labs(x = "Black to White Ratio", y = "Fatal Shootings Proportions", title = "Proportion of Fatal Shootings vs. Black to White Ratios in Cities") +
   theme(legend.text = element_text(size = 8),
@@
 ```
-Means are not that different, will probably just use percent black and percent white then.
+
+Can't really tell if there's a distribution here, maybe negatively correlated.
+
+```{r black percent}
+temp <- final_df %>% 
+  group_by(city) %>% 
+  summarise(bprop = black/city_pop_2013, propfatal = sum(fatal)/n()) %>% 
+  unique
+
+plot <- ggplot(temp, aes(x = bprop, y = propfatal)) + 
+    geom_point()  + 
+  labs(x = "Black Percentage of Population", y = "Fatal Shootings Proportions", title = "Proportion of Fatal Shootings vs. Black Population Percentages in Cities") +
+  theme(legend.text = element_text(size = 8),
+        legend.key.size = unit(0.4, "cm"),
+        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        plot.title = element_text(size=10, hjust = 0.5))
+
+ggsave(filename = "figures/bpropeda.png", plot)
+```
+
+Same thing, can't really tell if there's a distribution here, maybe negatively correlated.
+
+```{r}
+temp <- final_df %>% 
+  summarise(Black = sum(str_count(OfficerRace, "B")), 
+            White = sum(str_count(OfficerRace, "W")),
+            Hispanic = sum(str_count(OfficerRace, "L")),
+            Asian = sum(str_count(OfficerRace, "A")), 
+            Other = sum(str_count(OfficerRace, "O")))
+ggplot(temp, aes(x = ))
+```
+
+```{r}
+temp <- final_df %>% 
+  filter(sb == 1) %>% 
+  group_by(fatal) %>% 
+  summarise(ob = sum(ob == 1),
+            ow = sum(ow == 1),
+            ol = sum(ol == 1),
+            oa = sum(oa == 1),
+            oo = sum(oo == 1)) %>% 
+  mutate(ob = ob/(ob + ow + ol + oa + oo), 
+         ow = ow/(ob + ow + ol + oa + oo),
+         ol = ol/(ol + ow + ol + oa + oo), 
+         oa = oa/(ob + ow + ol + oa + oo), 
+         oo = oo/(ob + ow + ol + oa + oo))
+temp <- rename(temp, Black = ob, White = ow, Latino = ol, Asian = oa, Other = oo)
+
+temp2 <- temp %>%
+    t() %>%
+    as.data.frame(stringsAsFactors = F) %>%
+    rownames_to_column("race") %>%
+    `rownames<-`(NULL)
+temp2 <- rename(temp2, "Non-Fatal" = 2, Fatal = 3)
+temp2 <- temp2[-1,]
+temp2 <- melt(temp2)
+
+officerplot <- ggplot(temp2, aes(x=race, y = value, fill = variable)) +
+  geom_bar(position = "dodge", stat="identity") + 
+  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions for Fatal vs. Non-Fatal Police Shootings Where Black Victims were Present") +
+  theme(legend.text = element_text(size = 7),
+        legend.key.size = unit(0.4, "cm"),
+        plot.title = element_text(size=10, hjust = 0.5)) + 
+  scale_fill_manual(values=c("#D48C7D", "#7DB2D4"))
+
+ggsave(filename = "figures/officerpropsb.png", officerplot)
+```
 

### 2020-10-19 police_modeling.Rmd
eda

--- 2020-10-14 police_modeling.Rmd
@@
 library(tidyverse)
+library(lmtest)
 ```
@@
 ```{r load-data}
-all <- read.csv("allshootingsdata.csv")
+final_df <- readRDS("data/final.rds")
 
-philly <- read.csv("philly_shootings.csv")
-cincinnati <- read.csv("cincinnati_shootings.csv")
+single_victims <- final_df %>% 
+  filter(NumberOfSubjects <= 1)
+
+mult_victims <- final_df %>% 
+  filter(NumberOfSubjects > 1)
 ```
 
-```{r clean-all}
-all <- all %>% 
-  select(-c("Date", "NatureOfStop", "Notes", "SubjectAge", "NumberOfShots", "Department", "FullNarrative", "fdate", "month", "day", "b", "l", "w", "a", "tr", "AvgAge", "VA", "F", "nshots", "sit"))
+```{r}
+fit1 <- glm(formula = fatal ~ SubjectRace, family = binomial, data=single_victims)
 
-all <- all %>% 
-  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")
+fit2 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013), family = binomial, data=single_victims)
+lrtest(fit2, fit1)
+summary(fit2)
 
-officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
-officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
-out <- which((officer_gender == officer_race) == FALSE)
-all <- all[-out,]
+# Victim Gender doesn't appear to be significant
+fit3 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + SubjectGender, family = binomial, data=single_victims)
+lrtest(fit3, fit2)
 
-all <- all %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ",")
+fit4 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob, family = binomial, data=single_victims)
+lrtest(fit3, fit4)
 
-officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
-officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
-out2 <- which((officer_gender2 == officer_race2) == FALSE)
-all <- all[-out2,]
+fit5 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow, family = binomial, data=single_victims)
+lrtest(fit5, fit4)
 
-all$OfficerRace <- trimws(all$OfficerRace)
-all$OfficerGender <- trimws(all$OfficerGender)
-all$Fatal <- trimws(all$Fatal)
+fit6 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol, family = binomial, data=single_victims)
+lrtest(fit5, fit6)
 
-all <- all %>% 
-  filter(SubjectRace != "U") %>% 
-  filter(OfficerRace != "U") %>% 
-  filter(OfficerRace != "Unknown") %>% 
-  filter(OfficerRace != "NA") %>% 
-  filter(OfficerRace != "") %>% 
-  filter(OfficerRace != "I") %>% 
-  filter(OfficerGender != "U") %>% 
-  filter(OfficerGender != "N") %>% 
-  filter(SubjectGender != "U") %>% 
-  filter(SubjectGender != "N") %>% 
-  filter(Fatal != "U")
+fit7 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
+lrtest(fit6, fit7)
 
-all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
-all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
-all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"
+fit8 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
+lrtest(fit8, fit7)
 
-all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
-all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
-all$OfficerGender[all$OfficerGender == "Male"] <- "M"
+fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + SubjectArmed, family = binomial, data=single_victims)
+lrtest(fit8, fit9)
 ```

### 2020-10-19 submissionii.tex
eda

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Police Shootings and Racism in America},
  pdfauthor={Connie Wu},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{Police Shootings and Racism in America}
\author{Connie Wu}
\date{10/17/2020}

\begin{document}
\maketitle

\hypertarget{i.-introduction}{%
\subsection{I. Introduction}\label{i.-introduction}}

~~~~~~~On May 25, 2020, George Floyd was publicly suffocated to death by
police officers in Minneapolis, Minnesota. This, along with the murder
of Breonna Taylor by Louisville police, incited waves of protests
against police brutality across the United States and increased the
spread of Black Lives Matter content on social media. The recent boost
in attention to the Black Lives Matter movement has once again brought
to light the issue of racism in America and its link to police
brutality, and more specifically, police use of deadly force. Several
studies have already found that race does play a part in who is targeted
and killed in police shootings. For example, in a recent VICE News
investigation about police shootings, it was found that ``Black people
were shot more often and at higher rates than people of any other
race.'' {[}1{]} Additionally, Edwards et. al performed a study regarding
the effects of age, race-ethnicity, and sex on the risk of being killed
by lethal force by law enforcement and similarly found that ``Black men
are about 2.5 times more likely to be killed by police over the life
course than are white men'' while ``Black women are about 1.4 times more
likely to be killed by police than are white women.'' {[}2{]} However,
even with these studies and their disturbing conclusions, a poll done by
AP-NORC in June 2020 found that still 39\% of Americans think that
police are not more likely to use lethal force against a Black person
than a White person. {[}3{]} Although this has decreased from an
overwhelming 51\% in June 2015, there is still a great deal of research
that needs to be done in this area to provide more statistical evidence
backing the relationship between racism and police use of deadly
force.\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}As
a result, I have decided to build off of VICE News' study and
investigate data on police shootings further to understand the roles
that the race of both victim and officer, as well as other factors such
as their genders, whether they are carrying a weapon, and the total
number of victims in the crime, play in fatal versus non-fatal police
shootings. Additionally, to account for the varying locations of the
homicides, I will be using a dataset found on Kaggle detailing the gun
provisions that are upheld by each state see how gun legislation affects
lethal vs.~non-lethal shootings. I will also add a predictor indicating
whether the state in which the homicide occurred requires de-escalation
training for police officers. Finally, I will explore how the race
demographics of each location relates to police use of deadly force.
This will allow me to better understand how racism has manifested itself
in America's police system and determine whether current attempts to
prevent police use of lethal force are effective or not.

\hypertarget{data}{%
\subsubsection{Data}\label{data}}

~~~~~~~As a basis for this study, I will be using the same dataset that
VICE News used. This dataset contains data on officer-involved shootings
from 47 of the largest local police departments in America, and more
specifically, ``information on 4,117 incidents and 4,400 subjects
{[}(victims){]} over seven years.'' {[}1{]} VICE News provides 34
variables in the dataset, including dates ranging from January 2010 to
September 2016, city, subject race, subject gender, office race, officer
gender, the type of weapon the subject was carrying, and whether the
shooting was fatal or not. Race was broken up into 5 categories: White
(non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A),
Latino (L), Other (O), and Unknown (U). Gender was broken up into 3
categories: Male (M), Female (F), and Unknown (U). Weapon type was
broken up into 6 values: ``gun'', ``knife'', ``replica'', ``other'',
``unknown'', and ``other''. Additionally, there were cases in which
multiple victims and/or multiple officers were present in the shooting.
Each of these scenarios was still represented within one row of the
dataset, but semi-colons were present in the respective victim and
officer columns, separating each individual's information from each
other.\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}Additionally,
to assess the effect of different types of legislation on fatal
vs.~non-fatal police shootings in various states, I will be using a
dataset from Kaggle containing 135 variables that detail whether a
certain gun provision is absent or present in a certain year and U.S.
state for 133 different gun provisions. The years range from 1991 to
2017, and the gun provisions address 14 categories, some of which are
dealer regulations, buyer regulations, prohibitions for high-risk gun
possession, background checks, ammunition regulations, possession
regulations, concealed carry permitting, assault weapons and
large-capacity magazines, child access prevention, gun trafficking, and
domestic violence. A 1 in the gun provision column represents a presence
of the law, and a 0 indicates an absence. I will also add a column to
this gun provision dataset indicating whether a state requires
de-escalation training based off of information reported by APM Report
in 2017, with a 1 meaning the state does require training and a 0
meaning the state does not. There were 16 states that required
de-escalation training as of November 2017. {[}4{]}\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}Finally,
to examine city demographics as a predictor, I will be using population
data from the 2013 American Community Survey that VICE News had already
cleaned and standardized to the shootings dataset, meaning the city
names can be matched up between the two datasets. VICE News most likely
provided only 2013 census data because it is the average year of all the
years represented in the shootings dataset. Furthermore, using census
data only from 2013 requires that we analyze this data under the
assumption that there is no drastic difference in population
demographics between 2010 and 2016. The census data includes 7
variables: the police department and the city it's located in, as well
as the city's Black, Asian, Hispanic, White, and overall total
population in 2013.

\hypertarget{ii.-methods}{%
\subsection{II. Methods}\label{ii.-methods}}

\hypertarget{data-processing}{%
\subsubsection{Data Processing}\label{data-processing}}

~~~~~~~Although the VICE News dataset was already relatively clean with
regards to the victim's data, there was still quite a bit of cleaning
that needed to be done for the race and gender data of the officers. To
clean these columns, I replaced all values that were not ``W'', ``B'',
``A'', ``O'', or ``U'' in the race column with the most informed guess
that I could make about what the values represented. For example,
``A/PI'' values were taken to represent Asian/Pacific Islander and thus
were replaced with ``A'', and ``A/W'' or other values with a ``/'' in
them were typically taken to represent multi-racial individuals and as a
result were replaced with ``O.'' Similar procedures were carried out in
the gender column for officers for values that were not ``M'', ``F'', or
``U''.\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}After
this initial cleaning, I created new columns in the same dataset that
represented whether any victims of race Black, White, Asian, or Other
were present in the crime, respectively, and if any male or female
victims were present in the crime, respectively. I also added new
columns to represent the same information for officers (for each race
and gender, whether each was present), and a new column to represent
whether any victim involved in the shooting was shot fatally.
Additionally, I made sure that each weapon type had its own column. All
of these added columns had values of either 0 or 1, with 0 representing
an absence of the variable and 1 a presence. Finally, I filtered out
2,079 rows that had only unknown and NA values for the fatality of the
shooting or the races or genders of the victims or officers, as these
rows did not provide sufficient information for my analysis, and I
selected only the columns I needed, such as whether any victim of the
shooting was shot fatally or not, the genders and races of the victims
and officers, and the weapons that the victims were carrying, if any.\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}For
the gun provisions dataset, I created a new column that sums up the
total number of laws listed in the dataset that each state had in 2013
(for the same reason that VICE News used 2013). Additionally, as I
mentioned earlier, I appended a column to this dataset representing
whether a state requires de-escalation training for police officers as
of November 2017. 2017 was used under the assumption that no major event
occurred between September 2016 and November 2017 that caused a sudden
increase in the number of states that require de-escalation training for
police officers. Finally, for the census data, the only cleaning that
needed to be done was extracting the state that each city was in from
the department column so that the census could be easily joined with the
cleaned shootings dataset by city name, as well as the gun provisions
dataset by abbreviated state name.\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}After
merging the three datasets, the final dataset consists of 2,031
observations.

\hypertarget{exploratory-data-analysis}{%
\subsubsection{Exploratory Data
Analysis}\label{exploratory-data-analysis}}

\begin{figure}
\includegraphics[width=0.25\linewidth]{figures/cityeda1} \includegraphics[width=0.25\linewidth]{figures/cityeda2} \includegraphics[width=0.25\linewidth]{figures/cityeda3} \includegraphics[width=0.25\linewidth]{figures/cityeda4} \caption{Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data }\label{fig:figures-city}
\end{figure}

~~~~~~Plotted above in \emph{Figure 1} are the race proportions of
shooting victims in the top 4 most popular cities in the dataset
(Chicago, Los Angeles, Houston, and Phoenix) vs.~the race proportions of
those cities' total populations in 2013. Because the bars representing
shootings in the Black race category are taller than the bars
representing city-wide population in all 4 plots, it is evident Black
people are disproportionately the victims of police shootings compared
to city race proportions. This can especially be seen in Houston, where
Black people make up less than 25\% of the city population yet more than
50\% of the victim population of police shootings.

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/singleeda1} \includegraphics[width=0.5\linewidth]{figures/singleeda2} \caption{Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings}\label{fig:figures-single}
\end{figure}

~~~~~~Additionally, there appears to be 1,906 incidents in the final
dataset in which there was only one victim, and 148 with multiple
victims. Among all shootings with only one victim, it appears in
\emph{Figure 2} that the victim is more often Black than some other
race. However, it seems that Black victims and most victims of the other
race types are more likely to be shot non-fatally than fatally, whereas
White victims seem more likely to be shot fatally. Finally, it is
important to note that the numbers of victims that are of race types
Asian and Other are relatively small. This means that we should proceed
with caution when evaluating the effects of each of these races on
whether a shooting is fatal or non-fatal.\\
\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}\hspace*{0.333em}Similarly,
Black victims make up the highest proportion of victims in shootings
with multiple victims. However, in these shootings, all race types are
more likely to be shot non-fatally than fatally, with Latinos having the
highest proportion of being shot fatally out of all other race types.
Additionally, it should be noted that there are no victims with race
type ``Other'' in shootings where there are multiple victims. These
graphs can be found as \emph{Figure 3} in the Appendix.

\hypertarget{model}{%
\subsubsection{Model}\label{model}}

~~~~~~~The predictors that I will use are subject race, subject gender,
officer race, officer gender, and the type of weapon the subject was
carrying. The response I will be modeling will be whether the shooting
was fatal or not.

\hypertarget{iii.-preliminary-results}{%
\subsection{III. Preliminary Results}\label{iii.-preliminary-results}}

\begin{itemize}
\tightlist
\item
  limitations: time series, also counting N;U\ldots{} combos as N even
  though some are unknown
\end{itemize}

\hypertarget{references}{%
\subsection{References}\label{references}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Vice
\item
  \url{https://www.pnas.org/content/pnas/116/34/16793.full.pdf}
\item
  \url{https://www.kpbs.org/news/2020/jun/18/sweeping-change-us-views-police-violence-new-poll-/}
\item
  \url{http://govred.com/blog/deescalation-training-state-requirements/}
\end{enumerate}

\hypertarget{appendix}{%
\subsection{Appendix}\label{appendix}}

\hypertarget{exploratory-data-analysis-1}{%
\subsubsection{Exploratory Data
Analysis}\label{exploratory-data-analysis-1}}

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/multeda1} \includegraphics[width=0.5\linewidth]{figures/multeda2} \caption{Counts (left) and Proportions (right) of Different Races Across Multiple Victim Shootings}\label{fig:figures-mult}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/btweda} \caption{Proportion of Fatal Shootings vs. Black to White Ratios in Cities in Dataset}\label{fig:figures-btw}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/bpropeda} \caption{Proportion of Fatal Shootings vs. Black Population Percentages in Cities in Dataset}\label{fig:figures-bprop}
\end{figure}

\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/officerpropsb} \caption{Race Proportions for Fatal vs. Non-Fatal Police Shootings Where At Least One Black Victim was Present}\label{fig:figures-officersbprop}
\end{figure}

\end{document}

### 2020-10-19 police_eda.Rmd
added last of eda i hope

@@
             Other = sum(str_count(OfficerRace, "O")))
-ggplot(temp, aes(x = ))
+
+temp2 <- temp %>%
+    t() %>%
+    as.data.frame(stringsAsFactors = F) %>%
+    rownames_to_column("race") %>%
+    `rownames<-`(NULL)
+temp2 <- rename(temp2, Count = 2) %>% 
+  mutate(total = Count/sum(Count))
+
+officerplot2 <- ggplot(temp2, aes(x=race, y = total)) +
+  geom_bar(position = "dodge", stat="identity", fill ="#D48C7D") + 
+  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions of Officers for All Shootings") +
+  theme(legend.text = element_text(size = 7),
+        legend.key.size = unit(0.4, "cm"),
+        plot.title = element_text(size=10, hjust = 0.5))
+ggsave(filename = "figures/officerprop.png", officerplot2)
 ```
 
-```{r}
+```{r sb-officer}
 temp <- final_df %>% 
@@
   geom_bar(position = "dodge", stat="identity") + 
-  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions for Fatal vs. Non-Fatal Police Shootings Where Black Victims were Present") +
+  labs(x = "Officer Race", y = "Proportion", title = "Race Proportions of Officers Present at Fatal vs. Non-Fatal Police Shootings Involving Black Victims") +
   theme(legend.text = element_text(size = 7),
@@
 ```
+
+Nothing too crazy happening here.
 

### 2020-10-19 submissionii.Rmd
added last of eda i hope

@@
 
-|       Plotted below in \emph{Figure 1} are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs.
+```{r, figures-city, fig.show="hold", out.width="25%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
+knitr::include_graphics("figures/cityeda1.png")
+knitr::include_graphics("figures/cityeda2.png")
+knitr::include_graphics("figures/cityeda3.png")
+knitr::include_graphics("figures/cityeda4.png")
+```
+|       Plotted above in \emph{Figure 1} are the race proportions of shooting victims in the top 4 most popular cities in the dataset (Chicago, Los Angeles, Houston, and Phoenix) vs.
 the race proportions of those cities' total populations in 2013.
@@
 This can especially be seen in Houston, where Black people make up less than 25% of the city population yet more than 50% of the victim population of police shootings.
-```{r, figures-city, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
-knitr::include_graphics("figures/cityeda1.png")
-knitr::include_graphics("figures/cityeda2.png")
-```
-
 ```{r num_rows, eval = F}
@@
 ```
-
 ```{r, figures-single, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
@@
 
-```{r, figures-btw, fig.show="hold", out.width="50%", fig.cap="City Black to White Ratios Distributions for Fatal vs. Non-Fatal Police Shootings"}
+```{r, figures-btw, fig.show="hold", out.width="50%", fig.cap="Proportion of Fatal Shootings vs.
+Black to White Ratios in Cities in Dataset (left) and vs.
+Black Population Percentages in Cities in Dataset (right)"}
 par(mar = c(2, 2, .1, .1))
 knitr::include_graphics("figures/btweda.png")
+knitr::include_graphics("figures/bpropeda.png")
 ```
 
-Means are not that different, will probably just use percent black and percent white then.
+```{r, figures-officersbprop, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Officers For All Shootings (left) and For Fatal vs.
+Non-Fatal Police Shootings Where At Least One Black Victim was Present (right) and "}
+par(mar = c(2, 2, .1, .1))
+knitr::include_graphics("figures/officerprop.png")
+knitr::include_graphics("figures/officerpropsb.png")
+```

### 2020-10-19 police_eda.Rmd
model writing

@@
         legend.key.size = unit(0.4, "cm"),
-        axis.text.x = element_text(angle = 10, hjust = 1, size = 8),
+        axis.text.x = element_text(size = 8),
         plot.title = element_text(size=10, hjust = 0.5)) + 

### 2020-10-19 submissionii.Rmd
model writing

@@
 
-```{r, figures-city, fig.show="hold", out.width="25%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
+```{r, figures-city, fig.show="hold", out.width="30%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
 knitr::include_graphics("figures/cityeda1.png")
@@
 
-|        The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
+|        Because I want to explore the relationship between the number of victims in a police shooting affects and whether the shooting is fatal or not, I decided to create two separate models: one for shootings with only one victim, and one for shootings with multiple victims.
+Both models are fixed effect logistic regression models because the range that the response variable can take in logistic regression is the most reasonable given that I am predicting whether a victim is shot fatally or not (labeled 1 and 0, respectively).
+From the Exploratory Data Analysis section, it is evident that victim race plays a significant role in whether the shooting is fatal or not.
+Thus, for both models, I started The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
 The response I will be modeling will be whether the shooting was fatal or not.

### 2020-10-19 police_modeling.Rmd
modeling

@@
 lrtest(fit8, fit9)
+summary(fit9)
 ```

### 2020-10-19 submissionii.Rmd
modeling

@@
 [3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
-|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as their genders, whether they are carrying a weapon, and the total number of victims in the crime, play in fatal versus non-fatal police shootings.
+|        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as whether they are carrying a weapon and the total number of victims in the crime, play in fatal versus non-fatal police shootings.
 Additionally, to account for the varying locations of the homicides, I will be using a dataset found on Kaggle detailing the gun provisions that are upheld by each state see how gun legislation affects lethal vs.
@@
 From the Exploratory Data Analysis section, it is evident that victim race plays a significant role in whether the shooting is fatal or not.
-Thus, for both models, I started The predictors that I will use are subject race, subject gender, officer race, officer gender, and the type of weapon the subject was carrying.
-The response I will be modeling will be whether the shooting was fatal or not.
-```{r}
-
-```
+Thus, for both models, I started with a simple logistic regression model using only victim race as a predictor.
+From there, I began adding the other predictors that I wanted to explore one at a time and kept an eye on the p-values of all predictors to make sure they all remained significant.
+Additionally, I performed likelihood ratio tests with each new model iteration to check that the added variable improves the older model.
+|        The final model for single-victim shootings is written out below:
+\begin{center}
+$$
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ muEDA_i + \beta_2\ sdEDA_i + \beta_3\ muEDA_i:sdEDA_i + \beta_4\ muTemp_i + \beta_5\ sdTemp_i + \beta_6\ muEMG_i
+$$
+$$
++ \beta_7\ sdEMG_i + \beta_8\ muACC\_Y_i + \beta_9\ sdACC\_Y_i + \beta_{10}\ muACC\_Y_i:muEDA_i + \beta_{11}\ muEDA_i:muTemp_i
+$$
+\end{center}
+|        The final model for multiple-victim shootings written out below:
+\begin{center}
+$$
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ muEDA_i + \beta_2\ sdEDA_i + \beta_3\ muEDA_i:sdEDA_i + \beta_4\ muTemp_i + \beta_5\ sdTemp_i + \beta_6\ muEMG_i
+$$
+$$
++ \beta_7\ sdEMG_i + \beta_8\ muACC\_Y_i + \beta_9\ sdACC\_Y_i + \beta_{10}\ muACC\_Y_i:muEDA_i + \beta_{11}\ muEDA_i:muTemp_i
+$$
+\end{center}
 
@@
 
+### Analysis Plan
+
+|        For sensitivity analysis, I plan to perform cross-validation on the dataset by holding out 10% of the data to use as a test set and training my models on the other 90% of the data as a way to assess prediction accuracy.
+
+### Strengths and Limitations
+|        There are several limitations to this study.
+For example, the VICE News dataset spans over the years 2010 to 2016, and neither of my models will be exploring time series data whatsoever.
+Thus, the aspect of time may be playing some sort of effect on my response variable that I am not accurately capturing in my models.
+Additionally, when cleaning the data, there were several "Unknown" values in the fatality column.
+Unknown values were assumed to represent "No" in this study, although this may not necessarily be true, to allow for more straightforward cleaning of the data and not lose too many observations in the data.
+Finally, there were less than 200 cases in which there were multiple victims in my final dataset, which is a relatively small sample size.
 - limitations: time series, also counting N;U... combos as N even though some are unknown

### 2020-10-19 police_modeling.Rmd
modling

--- 2020-10-19 police_modeling.Rmd
@@
 
-fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + SubjectArmed, family = binomial, data=single_victims)
+fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + gun, family = binomial, data=single_victims)
 lrtest(fit8, fit9)
+summary(fit9)
 ```
+
+```{r}
+mult1 <- glm(formula = fatal ~ sb + sw + sl + I(black/city_pop_2013) + ob + ow + ol + ow:sb + gun, family = binomial, data=mult_victims)
+summary(mult1)
+```

### 2020-10-19 submissionii.Rmd
modling

@@
 |        As a basis for this study, I will be using the same dataset that VICE News used.
-This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from January 2010 to September 2016, city, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
+This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from January 2010 to September 2016, 47 unique cities, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
 Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), Other (O), and Unknown (U).
@@
 |        The final model for single-victim shootings is written out below:
+```{r, eval = F}
+fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + SubjectArmed, family = binomial, data=single_victims)
+```
+
 \begin{center}
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ muEDA_i + \beta_2\ sdEDA_i + \beta_3\ muEDA_i:sdEDA_i + \beta_4\ muTemp_i + \beta_5\ sdTemp_i + \beta_6\ muEMG_i
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \sum_{j=1}^{4} \beta_{j} I(SubjectRace_i = SubjectRace[j]) +
 $$
 $$
-+ \beta_7\ sdEMG_i + \beta_8\ muACC\_Y_i + \beta_9\ sdACC\_Y_i + \beta_{10}\ muACC\_Y_i:muEDA_i + \beta_{11}\ muEDA_i:muTemp_i
+\sum_{j=5}^{8}  \beta_{j} black\_officer\_present_i * I(SubjectRace_i = SubjectRace[j-4]) + \beta_9\ (\frac{black_i}{city\_pop_i}) +
+$$
+$$
+\beta_{10}\ black\_officer\_present_i + \beta_{11}\ white\_officer\_present_i + \beta_{12}\ latino\_officer\_present_i + \beta_{13} victim\_armed\_gun
+$$
+$$
+SubjectRace = [\text{Black, Latino, Other, White}]
 $$
 \end{center}
-|        The final model for multiple-victim shootings written out below:
+|        The final model for multiple-victim shootings is written out below:
 \begin{center}
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ muEDA_i + \beta_2\ sdEDA_i + \beta_3\ muEDA_i:sdEDA_i + \beta_4\ muTemp_i + \beta_5\ sdTemp_i + \beta_6\ muEMG_i
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ black\_victim\_present_i + \beta_2\ white\_victim\_present_i +  \beta_3\ latino\_victmi\_present_i + \beta_4\ (\frac{black_i}{city\_pop_i}) +
 $$
 $$
-+ \beta_7\ sdEMG_i + \beta_8\ muACC\_Y_i + \beta_9\ sdACC\_Y_i + \beta_{10}\ muACC\_Y_i:muEDA_i + \beta_{11}\ muEDA_i:muTemp_i
+\beta_{5}\ black\_officer\_present_i + \beta_{6}\ white\_officer\_present_i + \beta_{7}\ latino\_officer\_present_i
+$$
+$$
++ \beta_{8} victim\_armed\_gun + \beta_{9} victim\_armed\_gun
 $$
@@
 ### Analysis Plan
-
+|        Preliminarily, it appears that none of the legislation variables were significant in predicting whether one was shot fatally or not, and thus were not included in either of the models. 
 |        For sensitivity analysis, I plan to perform cross-validation on the dataset by holding out 10% of the data to use as a test set and training my models on the other 90% of the data as a way to assess prediction accuracy.
+For model diagnostics, I will plot binned residual plots and ROC curves to assess how well my models fit the data.
 
@@
 Finally, there were less than 200 cases in which there were multiple victims in my final dataset, which is a relatively small sample size.
-- limitations: time series, also counting N;U... combos as N even though some are unknown
+Thus, it may be necessary to take the results from the multiple-victim model with a grain of salt.
 

### 2020-10-19 submissionii.Rmd
finishing up

@@
 [3] Although this has decreased from an overwhelming 51% in June 2015, there is still a great deal of research that needs to be done in this area to provide more statistical evidence backing the relationship between racism and police use of deadly force.
+
 |        As a result, I have decided to build off of VICE News’ study and investigate data on police shootings further to understand the roles that the race of both victim and officer, as well as other factors such as whether they are carrying a weapon and the total number of victims in the crime, play in fatal versus non-fatal police shootings.
@@
 Each of these scenarios was still represented within one row of the dataset, but semi-colons were present in the respective victim and officer columns, separating each individual's information from each other.
+
 |       Additionally, to assess the effect of different types of legislation on fatal vs.
@@
 [4]
+
 |       Finally, to examine city demographics as a predictor, I will be using population data from the 2013 American Community Survey that VICE News had already cleaned and standardized to the shootings dataset, meaning the city names can be matched up between the two datasets.
@@
 For example, "A/PI" values were taken to represent Asian/Pacific Islander and thus were replaced with "A", and "A/W" or other values with a "/" in them were typically taken to represent multi-racial individuals and as a result were replaced with "O." Similar procedures were carried out in the gender column for officers for values that were not "M", "F", or "U".
+
 |        After this initial cleaning, I created new columns in the same dataset that represented whether any victims of race Black, White, Asian, or Other were present in the crime, respectively, and if any male or female victims were present in the crime, respectively.
@@
 Finally, I filtered out 2,079 rows that had only unknown and NA values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
+
 |        For the gun provisions dataset, I created a new column that sums up the total number of laws listed in the dataset that each state had in 2013 (for the same reason that VICE News used 2013).
@@
 Finally, for the census data, the only cleaning that needed to be done was extracting the state that each city was in from the department column so that the census could be easily joined with the cleaned shootings dataset by city name, as well as the gun provisions dataset by abbreviated state name.
+
 |        After merging the three datasets, the final dataset consists of 2,031 observations.
@@
 ```
+
 ```{r, figures-single, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Single Victim Shootings"}
@@
 ```
+
+```{r, figures-mult, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Multiple Victim Shootings"}
+par(mar = c(2, 2, .1, .1))
+knitr::include_graphics("figures/multeda1.png")
+knitr::include_graphics("figures/multeda2.png")
+```
+
 |       Additionally, there appears to be 1,906 incidents in the final dataset in which there was only one victim, and 148 with multiple victims.
@@
 This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.
-|       Similarly, Black victims make up the highest proportion of victims in shootings with multiple victims.
+
+|       Similarly, as seen in \emph{Figure 3}, Black victims make up the highest proportion of victims in shootings with multiple victims.
 However, in these shootings, all race types are more likely to be shot non-fatally than fatally, with Latinos having the highest proportion of being shot fatally out of all other race types.
 Additionally, it should be noted that there are no victims with race type "Other" in shootings where there are multiple victims.
-These graphs can be found as \emph{Figure 3} in the Appendix.
 
@@
 Additionally, I performed likelihood ratio tests with each new model iteration to check that the added variable improves the older model.
+
 |        The final model for single-victim shootings is written out below:
@@
 $$
-\beta_{10}\ black\_officer\_present_i + \beta_{11}\ white\_officer\_present_i + \beta_{12}\ latino\_officer\_present_i + \beta_{13} victim\_armed\_gun
+\beta_{10}\ black\_officer\_present_i + \beta_{11}\ white\_officer\_present_i + \beta_{12}\ latino\_officer\_present_i + \beta_{13} \ victim\_armed\_gun_i
 $$
@@
 \end{center}
+
 |        The final model for multiple-victim shootings is written out below:
@@
 $$
-+ \beta_{8} victim\_armed\_gun + \beta_{9} victim\_armed\_gun
++ \beta_{8} \ victim\_armed\_gun_i + \beta_{9} \ white\_officer\_present_i*black\_victim\_present_i
 $$
@@
 ### Analysis Plan
+
 |        Preliminarily, it appears that none of the legislation variables were significant in predicting whether one was shot fatally or not, and thus were not included in either of the models.
+Additionally, victim race and officer race both are significant individually and as part of interaction terms, which will be interesting to quantify.
+In the next revision, I hope to analyze the coefficients of these predictors in terms of odds ratios in order to determine the relationships between each of these predictors and police shooting fatality.
+
 |        For sensitivity analysis, I plan to perform cross-validation on the dataset by holding out 10% of the data to use as a test set and training my models on the other 90% of the data as a way to assess prediction accuracy.
@@
 ### Strengths and Limitations
-|        There are several limitations to this study.
+
+|        One strength of my analysis is that I am considering many variables as predictors, from individual level data to contextual predictors, such as state gun legislation.
+This allows me to thoroughly explore and determine whether the relationships between a wide range of potential predictors and police shooting fatality in the context of police brutality are significant in any types of shootings (single vs.
+multiple victims, for example).
+There are several limitations to this study.
 For example, the VICE News dataset spans over the years 2010 to 2016, and neither of my models will be exploring time series data whatsoever.
 Thus, the aspect of time may be playing some sort of effect on my response variable that I am not accurately capturing in my models.
+There are also several other confounding factors that have not been accounted for, such as how well one can shoot at baseline or what training they have received with regards to shooting.
 Additionally, when cleaning the data, there were several "Unknown" values in the fatality column.
@@
 ### Exploratory Data Analysis
-```{r, figures-mult, fig.show="hold", out.width="50%", fig.cap="Counts (left) and Proportions (right) of Different Races Across Multiple Victim Shootings"}
-par(mar = c(2, 2, .1, .1))
-knitr::include_graphics("figures/multeda1.png")
-knitr::include_graphics("figures/multeda2.png")
-```
 

### 2020-10-19 police_eda.Rmd
modeling
finsihed up sub 2

@@
 all$OfficerGender <- gsub(" |/", "", toupper(all$OfficerGender))
+
+all$SubjectArmed[which(is.na(all$SubjectArmed))] <- "U"
 
@@
 Nothing too crazy happening here.
-
-### OLD CODE 
-```{r}
-all <- all %>% 
-  separate_rows(SubjectRace, SubjectGender, Fatal, sep = ";")
-
-officer_gender <- sapply(str_split(all$OfficerGender, ";"), length)
-officer_race <- sapply(str_split(all$OfficerRace, ";"), length)
-out <- which((officer_gender == officer_race) == FALSE)
-all <- all[-out,]
-
-all <- all %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ";") %>% 
-  separate_rows(OfficerRace, OfficerGender, sep = ",")
-
-officer_gender2 <- sapply(str_split(all$OfficerGender, "/"), length)
-officer_race2 <- sapply(str_split(all$OfficerRace, "/"), length)
-out2 <- which((officer_gender2 == officer_race2) == FALSE)
-all <- all[-out2,]
-
-all$OfficerRace <- trimws(all$OfficerRace)
-all$OfficerGender <- trimws(all$OfficerGender)
-all$Fatal <- trimws(all$Fatal)
-
-all <- all %>% 
-  filter(SubjectRace != "U") %>% 
-  filter(OfficerRace != "U") %>% 
-  filter(OfficerRace != "Unknown") %>% 
-  filter(OfficerRace != "NA") %>% 
-  filter(OfficerRace != "") %>% 
-  filter(OfficerRace != "I") %>% 
-  filter(OfficerGender != "U") %>% 
-  filter(OfficerGender != "N") %>% 
-  filter(SubjectGender != "U") %>% 
-  filter(SubjectGender != "N") %>% 
-  filter(Fatal != "U")
-
-all$OfficerRace[all$OfficerRace == "WHITE"] <- "W"
-all$OfficerRace[all$OfficerRace == "BLACK"] <- "B"
-all$OfficerRace[all$OfficerRace == "ASIAN"] <- "A"
-
-all$OfficerGender[all$OfficerGender == "FEMALE"] <- "F"
-all$OfficerGender[all$OfficerGender == "MALE"] <- "M"
-all$OfficerGender[all$OfficerGender == "Male"] <- "M"
-```

### 2020-10-19 submissionii.Rmd
finsihed up sub 2

@@
 library(tidyverse)
+library(knitr)
 ```
@@
 state for 133 different gun provisions.
-The years range from 1991 to 2017, and the gun provisions address 14 categories, some of which are dealer regulations, buyer regulations, prohibitions for high-risk gun possession, background checks, ammunition regulations, possession regulations, concealed carry permitting, assault weapons and large-capacity magazines, child access prevention, gun trafficking, and domestic violence.
+[4] The years range from 1991 to 2017, and the gun provisions address 14 categories, some of which are dealer regulations, buyer regulations, prohibitions for high-risk gun possession, background checks, ammunition regulations, possession regulations, concealed carry permitting, assault weapons and large-capacity magazines, child access prevention, gun trafficking, and domestic violence.
 A 1 in the gun provision column represents a presence of the law, and a 0 indicates an absence.
@@
 There were 16 states that required de-escalation training as of November 2017.
-[4]
+[5]
 
@@
 Thus, for both models, I started with a simple logistic regression model using only victim race as a predictor.
-From there, I began adding the other predictors that I wanted to explore one at a time and kept an eye on the p-values of all predictors to make sure they all remained significant.
-Additionally, I performed likelihood ratio tests with each new model iteration to check that the added variable improves the older model.
+From there, I began adding the other predictors that I wanted to explore one at a time and performed likelihood ratio tests with each new model iteration to make sure that the added variable improves the previous model.
+This means that the likelihood ratio test's p-value is less than 0.05.
 
@@
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ black\_victim\_present_i + \beta_2\ white\_victim\_present_i +  \beta_3\ latino\_victmi\_present_i + \beta_4\ (\frac{black_i}{city\_pop_i}) +
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ black\_victim\_present_i + \beta_2\ white\_victim\_present_i +  \beta_3\ latino\_victmi\_present_i +
 $$
 $$
-\beta_{5}\ black\_officer\_present_i + \beta_{6}\ white\_officer\_present_i + \beta_{7}\ latino\_officer\_present_i
+\beta_4\ (\frac{black_i}{city\_pop_i}) + \beta_{5}\ black\_officer\_present_i + \beta_{6}\ white\_officer\_present_i +
 $$
 $$
-+ \beta_{8} \ victim\_armed\_gun_i + \beta_{9} \ white\_officer\_present_i*black\_victim\_present_i
+ \beta_{7}\ latino\_officer\_present_i + \beta_{8} \ victim\_armed\_gun_i + \beta_{9} \ white\_officer\_present_i*black\_victim\_present_i
 $$
@@
 
-|        Preliminarily, it appears that none of the legislation variables were significant in predicting whether one was shot fatally or not, and thus were not included in either of the models.
-Additionally, victim race and officer race both are significant individually and as part of interaction terms, which will be interesting to quantify.
-In the next revision, I hope to analyze the coefficients of these predictors in terms of odds ratios in order to determine the relationships between each of these predictors and police shooting fatality.
+|        Preliminarily, it appears that none of the legislation variables were significant in predicting whether one was shot fatally or not, and thus these were not included in either of the models.
+Additionally, victim and officer race both are significant individually and as part of interaction terms, which will be interesting to quantify.
+As expected, the coefficient for subject race type Black is positive in both models, meaning Black victims are more likely to be shot fatally.
+In the next revision, when quantifying the effects of each of these predictors on police shooting fatality, I will be analyzing the coefficients in terms of odds ratios, since I used logistic regression models.
+I also will most likely change the baseline to be White for both victim and officer race, since I want to investigate racism against minorities specifically in the law enforcement system.
 
@@
 ## References
-1) Vice 
-2) https://www.pnas.org/content/pnas/116/34/16793.full.pdf
-3) https://www.kpbs.org/news/2020/jun/18/sweeping-change-us-views-police-violence-new-poll-/
-4) http://govred.com/blog/deescalation-training-state-requirements/
+
+[1] McCann, Allison, et al. “Police Shoot Far More People than Anyone Realized, a VICE News Investigation Reveals.” VICE News, VICE News, 12 Dec. 2017, news.vice.com/en_us/article/xwvv3a/shot-by-cops.
+
+[2] Edwards, Frank, et al. “Risk of Being Killed by Police Use of Force in the United States by Age, Race–Ethnicity, and Sex.” PNAS, PNAS, 20 Aug. 2019, www.pnas.org/content/116/34/16793.
+
+[3] Press, Associated.
+“Sweeping Change In US Views Of Police Violence, New Poll Finds.” KPBS Public Media, KPBS, 18 June 2020, www.kpbs.org/news/2020/jun/18/sweeping-change-us-views-police-violence-new-poll-/.
+
+[4] https://www.kaggle.com/jboysen/state-firearms
+
+[5] Gilbert, Curtis.
+“Most States Neglect Ordering Police to Learn De-Escalation Tactics to Avoid Shootings.” Not Trained to Not Kill | APM Reports, APM Reports, 16 Sept.
+2020, www.apmreports.org/story/2017/05/05/police-de-escalation-training.
+
+[6] DeGue, Sarah et al.
+“Deaths Due to Use of Lethal Force by Law Enforcement: Findings From the National Violent Death Reporting System, 17 U.S.
+States, 2009-2012.” American journal of preventive medicine vol.
+51,5 Suppl 3 (2016): S173-S187.
+doi:10.1016/j.amepre.2016.08.027
 
@@
 
-### Exploratory Data Analysis
-
-```{r, figures-btw, fig.show="hold", out.width="50%", fig.cap="Proportion of Fatal Shootings vs.
-Black to White Ratios in Cities in Dataset (left) and vs.
-Black Population Percentages in Cities in Dataset (right)"}
+```{r, figures-btw, out.width="50%", fig.cap="Proportion of Fatal Shootings vs. Black to White Ratios in Cities in Dataset (left) and vs. Black Population Percentages in Cities in Dataset (right)"}
 par(mar = c(2, 2, .1, .1))
@@
 
-```{r, figures-officersbprop, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Officers For All Shootings (left) and For Fatal vs.
+```{r, figures-officersbprop, out.width="50%", fig.cap="Race Proportions of Officers For All Shootings (left) and For Fatal vs.
 Non-Fatal Police Shootings Where At Least One Black Victim was Present (right) and "}
@@
 ```
+
+```{r single and mult}
+single_victims <- final_df %>% 
+  filter(NumberOfSubjects <= 1)
+
+mult_victims <- final_df %>% 
+  filter(NumberOfSubjects > 1)
+```
+
+#### Single-Victim Shooting
+```{r single fit}
+fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + gun, family = binomial, data=single_victims)
+kable(summary(fit9)$coefficients)
+```
+
+#### Multiple-Victim Shooting
+```{r mult fit}
+mult1 <- glm(formula = fatal ~ sb + sw + sl + I(black/city_pop_2013) + ob + ow + ol + ow:sb + gun, family = binomial, data=mult_victims)
+kable(summary(mult1)$coefficients)
+```

### 2020-11-08 police_eda.Rmd
data cleaned again

@@
 ## Load Data
-
-```{r}
-library(ggplot2)
-library(dplyr)
-library(tidyr)
-library(tidyverse)
-library(reshape2)
-```
 
@@
 all$OfficerRace <- gsub("Other|AI/AN|NA/W|Multi-Racial", "O", all$OfficerRace)
+all$OfficerRace <- gsub("NA", "O", all$OfficerRace)
 all$OfficerRace <- gsub("A/PI ", "A;", all$OfficerRace)

### 2020-11-08 police_modeling.Rmd
data cleaned again

@@
 
+# Combining Asian and Other so model will converge
+final_df$SubjectRace[final_df$SubjectRace == "A"] <- "O"
+final_df$OfficerRace[final_df$OfficerRace == "A"] <- "O"
+
+final_df <- final_df %>% 
+  mutate(so = as.numeric(str_count(final_df$SubjectRace, "O|A") > 0),
+         oo = as.numeric(str_count(final_df$OfficerRace, "O|A") > 0)) %>% 
+  select(-c(sa, oa))
+
 single_victims <- final_df %>% 
-  filter(NumberOfSubjects <= 1)
+  filter(NumberOfSubjects <= 1) %>% 
+  mutate(SubjectRace = relevel(as.factor(SubjectRace), "W"))
 

### 2020-11-08 submissioniii.Rmd
data cleaned again

--- 2020-10-19 submissionii.Rmd
@@
 author: "Connie Wu"
-date: "10/17/2020"
+date: "11/09/2020"
 output:
@@
 This dataset contains data on officer-involved shootings from 47 of the largest local police departments in America, and more specifically, “information on 4,117 incidents and 4,400 subjects [(victims)] over seven years.” [1] VICE News provides 34 variables in the dataset, including dates ranging from January 2010 to September 2016, 47 unique cities, subject race, subject gender, office race, officer gender, the type of weapon the subject was carrying, and whether the shooting was fatal or not.
-Race was broken up into 5 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), Other (O), and Unknown (U).
+Race was broken up into 6 categories: White (non-Hispanic) (represented as W), Black (non-Hispanic) (B), Asian (A), Latino (L), Other (O), and Unknown (U).
 Gender was broken up into 3 categories: Male (M), Female (F), and Unknown (U).
-Weapon type was broken up into 6 values: "gun", "knife", "replica", "other", "unknown", and "other".
+Weapon type was broken up into 5 values: "gun", "knife", "replica", "other", and "unknown".
 Additionally, there were cases in which multiple victims and/or multiple officers were present in the shooting.
@@
 
-|        After merging the three datasets, the final dataset consists of 2,031 observations.
+|        After merging the three datasets, the final dataset consists of 2,054 observations.
 
@@
 
+***Since A and O are so little in the single victims dataset, I will combine them.
+
 ### Model
@@
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ black\_victim\_present_i + \beta_2\ white\_victim\_present_i +  \beta_3\ latino\_victmi\_present_i +
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ black\_victim\_present_i + \beta_2\ white\_victim\_present_i +  \beta_3\ latino\_victim\_present_i +
 $$
@@
 ## III. Preliminary Results
+
+TO DO: 
+- fix my models -- backward selection 
+- random effects by city, maybe race | city 
+- combine other with asian? 
+- more interactions
 

### 2020-11-08 submissioniii.Rmd
data cleaned again

@@
 All of these added columns had values of either 0 or 1, with 0 representing an absence of the variable and 1 a presence.
-Finally, I filtered out 2,079 rows that had only unknown and NA values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
+Finally, I filtered out 2,063 rows that had only unknown and NA values for the fatality of the shooting or the races or genders of the victims or officers, as these rows did not provide sufficient information for my analysis, and I selected only the columns I needed, such as whether any victim of the shooting was shot fatally or not, the genders and races of the victims and officers, and the weapons that the victims were carrying, if any.
 

### 2020-11-09 police_eda.Rmd
results

@@
 
+numofficerna <- which(is.na(all$NumberOfOfficers))
+replacement <- str_count(all$OfficerRace[numofficerna], ";")
+replacement <- replacement + 1
+all$NumberOfOfficers[numofficerna] <- replacement
+
 all <- all %>% 
@@
          ofem = as.numeric(str_count(all$OfficerGender, "F") > 0),
-         other = as.numeric(other), 
-         replica = as.numeric(replica), 
-         knife = as.numeric(knife), 
+         omult = as.numeric(all$NumberOfOfficers != "1"),
+         other = as.numeric(weapon == "other"), 
+         replica = as.numeric(weapon == "replica"), 
+         knife = as.numeric(weapon == "knife"), 
          gun = as.numeric(weapon == "gun"), 
@@
 final_df$NumberOfSubjects[which(final_df$NumberOfSubjects == 0)] <- 1
+final_df$NumberOfOfficers[which(final_df$NumberOfOfficers == "U")] <- 1
 ```

### 2020-11-09 police_modeling.Rmd
results

@@
 library(lmtest)
+library(lme4)
+library(caret)
+library(groupdata2) # fold()
 ```
@@
 
+### Old Validation Code
+
 ```{r}
-fit1 <- glm(formula = fatal ~ SubjectRace, family = binomial, data=single_victims)
+crossvalidate <- function(data, k, model, dependent, random = FALSE){
+  # 'data' is the training set with the ".folds" column
+  # 'k' is the number of folds we have
+  # 'model' is a string describing a linear regression model formula
+  # 'dependent' is a string with the name of the score column we want to predict
+  # 'random' is a logical; do we have random effects in the model?
 
-fit2 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013), family = binomial, data=single_victims)
-lrtest(fit2, fit1)
-summary(fit2)
+  # Initialize empty list for recording performances
+  performances <- c()
 
-# Victim Gender doesn't appear to be significant
-fit3 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + SubjectGender, family = binomial, data=single_victims)
-lrtest(fit3, fit2)
+  # One iteration per fold
+  for (fold in 1:k){
 
-fit4 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob, family = binomial, data=single_victims)
-lrtest(fit3, fit4)
+    # Create training set for this iteration
+    # Subset all the datapoints where .folds does not match the current fold
+    training_set <- data[data$.folds != fold,]
 
-fit5 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow, family = binomial, data=single_victims)
-lrtest(fit5, fit4)
+    # Create test set for this iteration
+    # Subset all the datapoints where .folds matches the current fold
+    testing_set <- data[data$.folds == fold,]
 
-fit6 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol, family = binomial, data=single_victims)
-lrtest(fit5, fit6)
+    ## Train model
 
-fit7 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
-lrtest(fit6, fit7)
+    # If there is a random effect,
+    # use glmer() to train model
+    # else use glm()
 
-fit8 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace, family = binomial, data=single_victims)
-lrtest(fit8, fit7)
+    if (isTRUE(random)){
+      # Train linear mixed effects model on training set
+      m <- glmer(model, training_set, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+    } else {
+      # Train linear model on training set
+      m <- glm(model, training_set, family = "binomial", control=glm.control(maxit=50))
+    }
 
-fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + gun, family = binomial, data=single_victims)
-lrtest(fit8, fit9)
-summary(fit9)
+    ## Test model
+
+    # Predict the dependent variable in the testing_set with the trained model
+    predicted <- predict(m, testing_set, type="response", allow.new.levels = TRUE)
+    fit.convert <- ifelse(predicted < 0.5, 0, 1)
+    table(fit.convert, testing_set[[dependent]])
+    performances[fold] <- mean(fit.convert == testing_set[[dependent]])
+  }
+
+  return(performances)
+}
 ```
@@
 ```{r}
-mult1 <- glm(formula = fatal ~ sb + sw + sl + I(black/city_pop_2013) + ob + ow + ol + ow:sb + gun, family = binomial, data=mult_victims)
-summary(mult1)
+getAccuracy <- function(v) {
+  if (length(v) == 4) {
+      return(((v[1] + v[4])/sum(v)) * 100)
+  }
+  if (length(v) == 2) {
+    return((v[1]/sum(v)) * 100)
+  }
+}
 ```
+
+## Single Victims
+
+### Backwards Selection
+```{r backwards-selection}
+fullmodel <- glm(fatal ~ SubjectRace + SubjectGender + ob + ow + ol + oo + omale + ofem + omult + gun + weapon_other + replica + knife + unarmed + I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train, family = binomial, data=single_victims)
+step(fullmodel, direction = "backward")
+```
+
+Backwards selection gives: glm(formula = fatal ~ SubjectRace + SubjectGender + ob + ow + 
+    omult + gun + knife + unarmed + I(black/city_pop_2013), family = binomial, 
+    data = single_victims)
+
+```{r backwards-selection}
+fullmodel <- glm(fatal ~ SubjectRace + SubjectGender + ob + ow + ol + oo + omale + ofem + omult + gun + weapon_other + replica + knife + unarmed + I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train + city, family = binomial, data=single_victims)
+step(fullmodel, direction = "backward")
+```
+
+When I put city in, it turns out to be in the backwards model too -- should probably do random effects with city. Formula: glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + city, family = binomial, data = single_victims)
+
+### Trying the models
+
+```{r backwards-fit}
+fitback <- glm(formula = fatal ~ SubjectRace + SubjectGender + ob + ow + 
+    omult + gun + knife + unarmed + I(black/city_pop_2013), family = binomial, 
+    data = single_victims)
+summary(fitback)
+```
+
+```{r validate fitback}
+set.seed(1)
+Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)
+
+mod_fitback <- train(as.factor(fatal) ~ SubjectRace + SubjectGender + ob + ow + 
+    omult + gun + knife + unarmed + I(black/city_pop_2013),  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
+    )
+
+pred <- predict(mod_fitback, newdata=testing )
+confusionMatrix(data=pred, as.factor(testing$fatal))
+```
+
+```{r validate fitback again}
+set.seed(1)
+
+Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ SubjectRace + SubjectGender + ob + ow + omult + gun + knife + unarmed + I(black/city_pop_2013)'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = FALSE)
+p0
+mean(p0)
+
+back2 <- glm(m0, training, family = "binomial")
+preds <- predict(rand_fit1, testing, type = "response")
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+table(fit.convert, testing[['fatal']])
+mean(fit.convert == testing[['fatal']])
+```
+
+```{r fit with city}
+fitback1 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + city, family = binomial, data = single_victims)
+summary(fitback1)
+
+back2 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + city + city:SubjectRace, family = binomial, data = single_victims)
+lrtest(fitback1, back2)
+
+back3 <- glm(formula = fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + city + city:gun, family = binomial, data = single_victims)
+lrtest(fitback1, back3)
+```
+
+Interactions between city and race lead to better fit, makes it seem like I should vary race by city perhaps
+
+```{r validate fitback1}
+set.seed(1)
+Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)
+
+mod_fitback1 <- train(as.factor(fatal) ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + city,  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
+    )
+
+pred <- predict(mod_fitback1, newdata=testing )
+confusionMatrix(data=pred, as.factor(testing$fatal))
+```
+```{r validate back2}
+set.seed(1)
+Train <- createDataPartition(single_victims$fatal, p=0.9, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+ctrl <- trainControl(method = "repeatedcv", number = 5, savePredictions = TRUE)
+
+mod_back2 <- train(as.factor(fatal) ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + city + city:SubjectRace,  data=single_victims, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, control=glm.control(maxit=50)
+    )
+
+pred <- predict(mod_back2, newdata=testing )
+confusionMatrix(data=pred, as.factor(testing$fatal))
+```
+
+```{r random-effects}
+rand_fit1 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + (1 | city), data = single_victims, family = "binomial", nAGQ=50)
+summary(rand_fit1)
+coef(rand_fit1)
+fixef(rand_fit1)
+ranef(rand_fit1)
+
+classDFrand_fit1 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit1),0))
+rand_fit1pred <- xtabs(~ predicted + response, data = classDFrand_fit1)
+getAccuracy(rand_fit1pred)
+```
+
+```{r validate rand_fit1}
+set.seed(1)
+
+Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 | city)'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
+p0
+mean(p0)
+
+rand_fit1 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+preds <- predict(rand_fit1, testing, type = "response")
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+table(fit.convert, testing[['fatal']])
+mean(fit.convert == testing[['fatal']])
+```
+
+```{r vary subject race by city}
+rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+summary(rand_fit2)
+coef(rand_fit2)
+fixef(rand_fit2)
+ranef(rand_fit2)
+
+classDFrand_fit2 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit2),0))
+rand_fit2pred <- xtabs(~ predicted + response, data = classDFrand_fit2)
+getAccuracy(rand_fit2pred)
+```
+
+```{r validate rand_fit2}
+set.seed(1)
+
+Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (SubjectRace | city)'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
+p0
+mean(p0)
+
+rand_fit2 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+preds <- predict(rand_fit2, testing, type = "response")
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+table(fit.convert, testing[['fatal']])
+mean(fit.convert == testing[['fatal']])
+```
+
+```{r vary subject race by city and make it vary intercept too}
+rand_fit3 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
+    replica + knife + unarmed + (1 + SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+summary(rand_fit3)
+coef(rand_fit3)
+fixef(rand_fit3)
+ranef(rand_fit3)
+
+classDFrand_fit3 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit3),0))
+rand_fit3pred <- xtabs(~ predicted + response, data = classDFrand_fit3)
+getAccuracy(rand_fit3pred)
+```
+
+```{r validate rand_fit3}
+set.seed(1)
+
+Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + SubjectRace | city)'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
+p0
+mean(p0)
+
+rand_fit3 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+preds <- predict(rand_fit3, testing, type = "response")
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+table(fit.convert, testing[['fatal']])
+mean(fit.convert == testing[['fatal']])
+```
+
+```{r random effects 4}
+rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+summary(rand_fit4)
+coef(rand_fit4)
+fixef(rand_fit4)
+ranef(rand_fit4)
+
+classDFrand_fit4 <- data.frame(response = single_victims$fatal, predicted = round(fitted(rand_fit4),0))
+rand_fit4pred <- xtabs(~ predicted + response, data = classDFrand_fit4)
+getAccuracy(rand_fit4pred)
+```
+
+```{r validate rand_fit4}
+set.seed(1)
+
+Train <- createDataPartition(single_victims$fatal, p=0.85, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
+p0
+mean(p0)
+
+rand_fit3 <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+preds <- predict(rand_fit3, testing, type = "response")
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+table(fit.convert, testing[['fatal']])
+mean(fit.convert == testing[['fatal']])
+```
+
+Single victims final model!!
+
+## Multiple Victims
+```{r backwards-selection}
+fullmodel <- glm(fatal ~ sb  + sw + sl + so + sfem + smale + weapon_other + replica + knife + ob + ow + ol + oo + omale + ofem + omult + gun + unarmed + unknown_weapon +  I(black/city_pop_2013) + I(black/white) + total_pr + deesc_train, family = binomial, data=mult_victims)
+step(fullmodel, direction = "backward")
+```

### 2020-11-09 submissioniii.Rmd
results

@@
 
-### Data
+## II. Data
 
@@
 The census data includes 7 variables: the police department and the city it's located in, as well as the city's Black, Asian, Hispanic, White, and overall total population in 2013.
-
-## II. Methods
 
@@
 Finally, it is important to note that the numbers of victims that are of race types Asian and Other are relatively small.
-This means that we should proceed with caution when evaluating the effects of each of these races on whether a shooting is fatal or non-fatal.
+Because of this, we will combine Asian and Other race types into one single race category called "Other" for both officers and victims for our models.
 
@@
 
-***Since A and O are so little in the single victims dataset, I will combine them.
-
-### Model
+## III. Methods
 
 |        Because I want to explore the relationship between the number of victims in a police shooting affects and whether the shooting is fatal or not, I decided to create two separate models: one for shootings with only one victim, and one for shootings with multiple victims.
-Both models are fixed effect logistic regression models because the range that the response variable can take in logistic regression is the most reasonable given that I am predicting whether a victim is shot fatally or not (labeled 1 and 0, respectively).
-From the Exploratory Data Analysis section, it is evident that victim race plays a significant role in whether the shooting is fatal or not.
-Thus, for both models, I started with a simple logistic regression model using only victim race as a predictor.
-From there, I began adding the other predictors that I wanted to explore one at a time and performed likelihood ratio tests with each new model iteration to make sure that the added variable improves the previous model.
-This means that the likelihood ratio test's p-value is less than 0.05.
+Both models are random effect logistic regression models because the range that the response variable can take in logistic regression is the most reasonable given that I am predicting whether a victim is shot fatally or not (labeled 1 and 0, respectively).
+Additionally, I will be using random effects to control for correlation within cities.
+
+|        To create these two models, I first performed backwards selection using simple logistic regression models on all relevant predictors in order to get a sense of which predictors are most statistically significant in prediction shooting fatality.
+Then, I added potential interactions that I wanted to explore one at a time and performed likelihood ratio tests with each new model iteration to check if the added interaction improved the previous model.
+Whether an interaction was statistically significant or not was determined by whether the likelihood ratio test's p-value was less than 0.05.
+Finally, I put these variables obtained from backwards selection into a random effects linear regression model with city as the random intercept and evaluated the fit of this model using cross validation, which I will describe in more detail in the Results section.
+From there, I tweaked the model by experimenting with additional predictors, interactions, and random effects between city and other predictors in order to obtain the best fitting model.
 
@@
 ```{r, eval = F}
-fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + SubjectArmed, family = binomial, data=single_victims)
+rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 ```
@@
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \sum_{j=1}^{4} \beta_{j} I(SubjectRace_i = SubjectRace[j]) +
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ SubjectRace_i + \beta_2\ OfficerBlack_i + \beta_3\ OfficerWhite_i + \beta_4\ MultipleOfficers_i
 $$
 $$
-\sum_{j=5}^{8}  \beta_{j} black\_officer\_present_i * I(SubjectRace_i = SubjectRace[j-4]) + \beta_9\ (\frac{black_i}{city\_pop_i}) +
++ \beta_{5j}\ VictimWithGun_{ij} + \beta_6\ VictimWithOther_i + \beta_7\ VictimWithReplica_i + \beta_8\ VictimWithKnife_i + \beta_9\ VictimUnarmed_i + b_i
 $$
 $$
-\beta_{10}\ black\_officer\_present_i + \beta_{11}\ white\_officer\_present_i + \beta_{12}\ latino\_officer\_present_i + \beta_{13} \ victim\_armed\_gun_i
+b_i \sim N(0, \sigma_b^2)
 $$
 $$
-SubjectRace = [\text{Black, Latino, Other, White}]
+SubjectRace = \text{{White*, Black, Latino, Other}}
 $$
 \end{center}
+In the model above, it is important to note that the predictor for whether the victim was carrying a gun is varied by city.
+This means that there is both a random intercept for city as well as a random slope for whether the victim was carrying a gun.
 
@@
 
-## III. Preliminary Results
+## IV. Results
 
+### Single Victims Model
+
+```{r random effects 4}
+rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+summary(rand_fit4)
+coef(rand_fit4)
+fixef(rand_fit4)
+ranef(rand_fit4)
+```
+
+### Multiple Victims Model
+
+### Model Validation and Diagnostics
+
+### Sensitivity Analysis
 TO DO: 
@@
 - random effects by city, maybe race | city 
-- combine other with asian? 
 - more interactions
@@
 For model diagnostics, I will plot binned residual plots and ROC curves to assess how well my models fit the data.
+
+## V. Discussion and Conclusion
 
@@
 
-## References
+## VI. References
 
@@
 
-## Appendix
+## VII. Appendix
 

### 2020-11-09 police_modeling.Rmd
results

@@
 rand_fit1 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
-    replica + knife + unarmed + (1 | city), data = single_victims, family = "binomial", nAGQ=50)
+    replica + knife + (1 | city), data = single_victims, family = "binomial", nAGQ=50)
 summary(rand_fit1)
@@
 
-m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 | city)'
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (1 | city)'
 p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
@@
 ```{r vary subject race by city}
-rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
-    replica + knife + unarmed + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 summary(rand_fit2)
@@
 
-m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (SubjectRace | city)'
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (SubjectRace | city)'
 p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
@@
 ```
+Single victims final model!!
 
@@
 rand_fit3 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + 
-    replica + knife + unarmed + (1 + SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+    replica + knife + (1 + SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 summary(rand_fit3)
@@
 
-m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + SubjectRace | city)'
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (1 + SubjectRace | city)'
 p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
@@
 ```{r random effects 4}
-rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 summary(rand_fit4)
@@
 
-m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender'
+m0 <- 'fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (1 + gun | city) + SubjectGender'
 p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
@@
 ```
+## Multiple Victims
 
-Single victims final model!!
-
-## Multiple Victims
 ```{r backwards-selection}
@@
 ```
+
+```{r mult}
+rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica +  (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+summary(rand_fitm)
+coef(rand_fitm)
+fixef(rand_fitm)
+ranef(rand_fitm)
+
+classDFrand_fitm <- data.frame(response = mult_victims$fatal, predicted = round(fitted(rand_fitm),0))
+rand_fitmpred <- xtabs(~ predicted + response, data = classDFrand_fitm)
+getAccuracy(rand_fitmpred)
+```
+
+```{r validate rand_fit2}
+set.seed(1)
+
+Train <- createDataPartition(mult_victims$fatal, p=0.85, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city)'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
+p0
+mean(p0)
+
+rand_fitm <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+preds <- predict(rand_fitm, testing, type = "response", allow.new.levels=TRUE)
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+table(fit.convert, testing[['fatal']])
+mean(fit.convert == testing[['fatal']])
+```

### 2020-11-09 submissioniii.Rmd
results

@@
 library(knitr)
+library(lme4)
+library(caret)
+library(groupdata2)
 ```
@@
 final_df <- readRDS("data/final.rds")
+
+# Combining Asian and Other so model will converge
+final_df$SubjectRace[final_df$SubjectRace == "A"] <- "O"
+final_df$OfficerRace[final_df$OfficerRace == "A"] <- "O"
+
+final_df <- final_df %>% 
+  mutate(so = as.numeric(str_count(final_df$SubjectRace, "O|A") > 0),
+         oo = as.numeric(str_count(final_df$OfficerRace, "O|A") > 0)) %>% 
+  select(-c(sa, oa))
+
+single_victims <- final_df %>% 
+  filter(NumberOfSubjects <= 1) %>% 
+  mutate(SubjectRace = relevel(as.factor(SubjectRace), "W"))
+
+mult_victims <- final_df %>% 
+  filter(NumberOfSubjects > 1)
+```
+
+```{r}
+crossvalidate <- function(data, k, model, dependent, random = FALSE){
+  # 'data' is the training set with the ".folds" column
+  # 'k' is the number of folds we have
+  # 'model' is a string describing a linear regression model formula
+  # 'dependent' is a string with the name of the score column we want to predict
+  # 'random' is a logical; do we have random effects in the model?
+
+  # Initialize empty list for recording performances
+  performances <- c()
+
+  # One iteration per fold
+  for (fold in 1:k){
+
+    # Create training set for this iteration
+    # Subset all the datapoints where .folds does not match the current fold
+    training_set <- data[data$.folds != fold,]
+
+    # Create test set for this iteration
+    # Subset all the datapoints where .folds matches the current fold
+    testing_set <- data[data$.folds == fold,]
+
+    ## Train model
+
+    # If there is a random effect,
+    # use glmer() to train model
+    # else use glm()
+
+    if (isTRUE(random)){
+      # Train linear mixed effects model on training set
+      m <- glmer(model, training_set, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+    } else {
+      # Train linear model on training set
+      m <- glm(model, training_set, family = "binomial", control=glm.control(maxit=50))
+    }
+
+    ## Test model
+
+    # Predict the dependent variable in the testing_set with the trained model
+    predicted <- predict(m, testing_set, type="response", allow.new.levels = TRUE)
+    fit.convert <- ifelse(predicted < 0.5, 0, 1)
+    table(fit.convert, testing_set[[dependent]])
+    performances[fold] <- mean(fit.convert == testing_set[[dependent]])
+  }
+
+  return(performances)
+}
+```
+
+```{r}
+getAccuracy <- function(v) {
+  if (length(v) == 4) {
+      return(((v[1] + v[4])/sum(v)) * 100)
+  }
+  if (length(v) == 2) {
+    return((v[1]/sum(v)) * 100)
+  }
+}
 ```
@@
 
-```{r, figures-city, fig.show="hold", out.width="30%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data "}
+```{r, figures-city1, fig.show="hold", out.width="50%"}
+par(mar = c(2, 2, .1, .1))
 knitr::include_graphics("figures/cityeda1.png")
 knitr::include_graphics("figures/cityeda2.png")
+```
+
+```{r, figures-city2, fig.show="hold", out.width="50%", fig.cap="Race Proportions of Shooting Victims vs. City Population Demographics of the Most Popular 4 Cities in the Data"}
+par(mar = c(2, 2, .1, .1))
 knitr::include_graphics("figures/cityeda3.png")
@@
 
-|        The final model for single-victim shootings is written out below:
+|        The final model for single-victim shootings is written out below (where $p_i$ represents the probability of getting fatally shot for shooting $i$):
 ```{r, eval = F}
-rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 ```
@@
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ SubjectRace_i + \beta_2\ OfficerBlack_i + \beta_3\ OfficerWhite_i + \beta_4\ MultipleOfficers_i
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_{1j}\ VictimRace_{ij} + \beta_2\ OfficerBlack_i + \beta_3\ OfficerWhite_i + \beta_4\ MultipleOfficers_i
 $$
 $$
-+ \beta_{5j}\ VictimWithGun_{ij} + \beta_6\ VictimWithOther_i + \beta_7\ VictimWithReplica_i + \beta_8\ VictimWithKnife_i + \beta_9\ VictimUnarmed_i + b_i
++ \beta_5\ VictimWithGun_i + \beta_6\ VictimWithOther_i + \beta_7\ VictimWithReplica_i + \beta_8\ VictimWithKnife_i
+$$
+$$
+ + \beta_9\ VictimUnarmed_i + b_i
 $$
@@
 $$
-SubjectRace = \text{{White*, Black, Latino, Other}}
+VictimRace = \text{{White*, Black, Latino, Other}}
 $$
 \end{center}
-In the model above, it is important to note that the predictor for whether the victim was carrying a gun is varied by city.
-This means that there is both a random intercept for city as well as a random slope for whether the victim was carrying a gun.
+|        For single victims, the best fitting model was the one that used city as a random intercept and had random slopes for the SubjectRace in each of these cities.
 
-|        The final model for multiple-victim shootings is written out below:
+|        The final model for multiple-victim shootings is written out below (where $p_i$ represents the probability of getting fatally shot for shooting $i$):
+
+```{r mult-fit, eval=F}
+rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+```
 \begin{center}
 $$
-log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_1\ black\_victim\_present_i + \beta_2\ white\_victim\_present_i +  \beta_3\ latino\_victim\_present_i +
+log(\frac{p_i}{(1-p_i)}) = \beta_0 + \beta_{1j}\ VictimBlack_{ij} + \beta_{2j}\ VictimLatino_{ij} + \beta_3\ VictimOther_i
 $$
 $$
-\beta_4\ (\frac{black_i}{city\_pop_i}) + \beta_{5}\ black\_officer\_present_i + \beta_{6}\ white\_officer\_present_i +
++ \beta_4\ OfficerBlack_i + \beta_5\ OfficerLatino_i + \beta_6\ MultipleOfficers_i + \beta_7\ VictimWithGun_i
 $$
 $$
- \beta_{7}\ latino\_officer\_present_i + \beta_{8} \ victim\_armed\_gun_i + \beta_{9} \ white\_officer\_present_i*black\_victim\_present_i
+\beta_8\ VictimWithOther_i + \beta_9\ VictimWithReplica_i + \beta_{10}\ VictimWithKnife_i + \beta_{11}\ (\frac{black_i}{white_i}) + b_i
+$$
+$$
+b_i \sim N(0, \sigma_b^2)
 $$
 \end{center}
+|        For multiple victims, the best fitting model was the one that used city as a random intercept and had random slopes for whether a Black victim was present and whether a Latino victim was present in the shootings in each of these cities.
+It is important to note that none of the legislative predictors were useful in predicting the odds of someone being fatally shot in police shootings in either model.
+Additionally, $\frac{black_i}{white_i}$ refers to the proportion of black to white people in a city.
+This proved to be useful in increasing the predictive accuracy of the model.
 
@@
 
-```{r random effects 4}
-rand_fit4 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + unarmed + (1 + gun | city) + SubjectGender, data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
-summary(rand_fit4)
-coef(rand_fit4)
-fixef(rand_fit4)
-ranef(rand_fit4)
+```{r single random effects}
+rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+kable(summary(rand_fit2)$coefficients)
 ```
+
+|        In this single victim model, the $\beta_0$ coefficient, which has a value of -0.57, represents the log odds of being fatally shot by police in a 'typical' scenario and city.
+'Typical' can be defined as a shooting in which the victim is White and unarmed and there is only one officer present, who is neither Black nor White.
+Thus, $\beta_0 + b_i$, or the city-specific intercept, represents the log odds of being fatally shot for a 'typical' White, unarmed victim in city $i$ when there is only one officer present, who is neither Black nor White.
+The variabilities of the intercepts and slopes across different cities and different race groups within cities are all greater than 0, indicating that there is enough variation in the baseline odds of being shot fatally in different cities and different race groups within cities.
+From the fixed effect coefficients, we can see that generally, all other race types are less likely to be shot fatally than White 'typical' individuals.
+
+|        The `ranef` function allows us to further quantify the difference between the average predicted log odds of being fatally shot for the general population versus that of a specific city and subject race type, other predictors remaining unchanged.
+This can be found in the Appendix.
+From this table, we can see that the city with the highest baseline odds of being shot fatally is Albuquerque, whereas the city with the lowest baseline odds of being shot fatally is Chicago.
+For White people in Chicago, the predicted log odds of being shot fatally is 0.76 less than that of in a 'typical' scenario and city.
+However, for Black people in Chicago, the predicted log odds of being shot fatally is actually 0.28 higher than that of a 'typical' scenario and city.
+In contrast, for Latino people in Chicago, the predicted log odds of being shot fatally is 0.23 lower than that of a 'typical' scenario and city, and for people of other race types in Chicago, the predicted log odds of being shot fatally is 2.86 lower than that of a 'typical' scenario and city.
+We can conclude that in Chicago, Black people are disproportionately shot fatally in police shootings.
+In Albuquerque, the trend is almost opposite: for Black people, the log odds of being shot fatally is 0.43 less than that of a 'typical' scenario and city, whereas for White people, the log odds of being shot fatally is 0.67 more.
+Overall, we can conclude that city plays a large role in whether a race type is fatally shot in police shootings, although the general trend is that White people are most likely to be shot fatally than other race types.
 
@@
 
+```{r mult random effects}
+rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+kable(summary(rand_fitm)$coefficients)
+kable(ranef(rand_fitm))
+```
+
+|        In this multiple victims model, the $\beta_0$ coefficient, which has a value of -1.77, represents the log odds of being fatally shot by police in a 'typical' scenario and city.
+'Typical' can be defined as a shooting in which the victim is White and unarmed and there is only one officer present, who is neither Black nor Latino.
+Thus, $\beta_0 + b_i$, or the city-specific intercept, represents the log odds of being fatally shot for a 'typical' White, unarmed victim in city $i$ when there is only one officer present, who is neither Black nor White.
+The variabilities of the intercepts and slopes across different cities and different race groups within cities are all greater than 0, indicating that there is enough variation in the baseline odds of being shot fatally in different cities and different race groups within cities.
+From the fixed effect coefficients, we can see that generally, all other race types are less likely to be shot fatally than White 'typical' individuals.
+
+|        The `ranef` function also allows us to quantify the difference between the average predicted log odds of being fatally shot for the general population versus that of a specific city and subject race type, other predictors remaining unchanged.
+This can be found in the Appendix.
+From this table, we can see that the city with the highest baseline odds of being shot fatally is Albuquerque, whereas the city with the lowest baseline odds of being shot fatally is Chicago.
+For White people in Chicago, the predicted log odds of being shot fatally is 0.76 less than that of in a 'typical' scenario and city.
+However, for Black people in Chicago, the predicted log odds of being shot fatally is actually 0.28 higher than that of a 'typical' scenario and city.
+In contrast, for Latino people in Chicago, the predicted log odds of being shot fatally is 0.23 lower than that of a 'typical' scenario and city, and for people of other race types in Chicago, the predicted log odds of being shot fatally is 2.86 lower than that of a 'typical' scenario and city.
+We can conclude that in Chicago, Black people are disproportionately shot fatally in police shootings.
+In Albuquerque, the trend is almost opposite: for Black people, the log odds of being shot fatally is 0.43 less than that of a 'typical' scenario and city, whereas for White people, the log odds of being shot fatally is 0.67 more.
+Overall, we can conclude that city plays a large role in whether a race type is fatally shot in police shootings, although the general trend is that White people are most likely to be shot fatally than other race types.
+
+### Differences Between Single and Multiple Victims
+
 ### Model Validation and Diagnostics
 
+|        As I mentioned earlier, I performed 5-fold cross-validation on my models to assess model accuracy and fit.
+I chose to use 5 folds and 85% of the data because this is commonly used in practice, but I tried it with 90% of the data as well just to verify that my cross-validation technique was not biased for the percentage of the full data allocated for the training set.
+I used the `groupdata2` R package to partition our data set so that the training set contained 85% of the data from the full dataset, and then I ran 5-fold cross-validation on this training set before predicting on the test set.
+When assessing the performance of different models, we considered a higher average cross-validation and test set prediction accuracy to be indicative of a better model.
+For the single victims model, I achieved an average 5-fold cross-validation prediction accuracy of 0.66 and a test set prediction accuracy of 0.70.
+For the multiple victims model, I achieved an average 5-fold cross-validation prediction accuracy of 0.61, and a test set prediction accuracy of 0.63.
+
 ### Sensitivity Analysis
-TO DO: 
-- fix my models -- backward selection 
-- random effects by city, maybe race | city 
-- more interactions
-
-### Analysis Plan
-
-|        Preliminarily, it appears that none of the legislation variables were significant in predicting whether one was shot fatally or not, and thus these were not included in either of the models.
-Additionally, victim and officer race both are significant individually and as part of interaction terms, which will be interesting to quantify.
-As expected, the coefficient for subject race type Black is positive in both models, meaning Black victims are more likely to be shot fatally.
-In the next revision, when quantifying the effects of each of these predictors on police shooting fatality, I will be analyzing the coefficients in terms of odds ratios, since I used logistic regression models.
-I also will most likely change the baseline to be White for both victim and officer race, since I want to investigate racism against minorities specifically in the law enforcement system.
-
-|        For sensitivity analysis, I plan to perform cross-validation on the dataset by holding out 10% of the data to use as a test set and training my models on the other 90% of the data as a way to assess prediction accuracy.
-For model diagnostics, I will plot binned residual plots and ROC curves to assess how well my models fit the data.
 
@@
 
-### Strengths and Limitations
-
+|        Overall, 
 |        One strength of my analysis is that I am considering many variables as predictors, from individual level data to contextual predictors, such as state gun legislation.
@@
 ```{r single fit}
-fit9 <- glm(formula = fatal ~ SubjectRace + I(black/city_pop_2013) + ob + ow + ol + ow:SubjectRace + gun, family = binomial, data=single_victims)
-kable(summary(fit9)$coefficients)
+rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+kable(ranef(rand_fit2), caption = "Random Effects Coefficients for Single-Victim Shootings")
 ```
@@
 ```{r mult fit}
-mult1 <- glm(formula = fatal ~ sb + sw + sl + I(black/city_pop_2013) + ob + ow + ol + ow:sb + gun, family = binomial, data=mult_victims)
-kable(summary(mult1)$coefficients)
+rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+kable(ranef(rand_fitm), caption ="Random Effects Coefficients for Multiple-Victim Shootings")
 ```
+```{r validate rand_fit2}
+set.seed(1)
+
+Train <- createDataPartition(mult_victims$fatal, p=0.85, list=FALSE)
+training <- single_victims[ Train, ]
+testing <- single_victims[ -Train, ]
+
+data <- fold(
+  training,
+  k = 5
+) %>%
+  arrange(.folds)
+
+m0 <- 'fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city)'
+p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
+paste0("Average cross-validation prediction accuracy: ", mean(p0))
+
+rand_fitm <- glmer(m0, training, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+preds <- predict(rand_fitm, testing, type = "response", allow.new.levels=TRUE)
+fit.convert <- ifelse(preds < 0.5, 0, 1)
+kable(table(fit.convert, testing[['fatal']]))
+m <- mean(fit.convert == testing[['fatal']])
+paste0("Test set prediction accuracy: ", m)
+```

### 2020-11-09 submissioniii.Rmd
results

@@
 
-## I. Introduction
+# I. Introduction
 
@@
 
-## II. Data
+# II. Data
 
@@
 
-### Data Processing
+## Data Processing
 
@@
 
-### Exploratory Data Analysis
+## Exploratory Data Analysis
 
@@
 
-## III. Methods
+# III. Methods
 
@@
 ```{r mult-fit, eval=F}
-rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 ```
@@
 $$
-\beta_8\ VictimWithOther_i + \beta_9\ VictimWithReplica_i + \beta_{10}\ VictimWithKnife_i + \beta_{11}\ (\frac{black_i}{white_i}) + b_i
+\beta_8\ VictimWithOther_i + \beta_9\ VictimWithReplica_i + \beta_{10}\ VictimWithKnife_i + b_i
 $$
@@
 It is important to note that none of the legislative predictors were useful in predicting the odds of someone being fatally shot in police shootings in either model.
-Additionally, $\frac{black_i}{white_i}$ refers to the proportion of black to white people in a city.
-This proved to be useful in increasing the predictive accuracy of the model.
 
-## IV. Results
+# IV. Results
 
-### Single Victims Model
+## Single Victims Model
 
@@
 
-|        In this single victim model, the $\beta_0$ coefficient, which has a value of -0.57, represents the log odds of being fatally shot by police in a 'typical' scenario and city.
-'Typical' can be defined as a shooting in which the victim is White and unarmed and there is only one officer present, who is neither Black nor White.
-Thus, $\beta_0 + b_i$, or the city-specific intercept, represents the log odds of being fatally shot for a 'typical' White, unarmed victim in city $i$ when there is only one officer present, who is neither Black nor White.
+|        In this single victim model, the $\beta_0$ coefficient, which has a value of -0.57, represents the log odds of being fatally shot by police in a 'typical' shooting.
+This can be defined as a shooting in which the victim is White and unarmed and there is only one officer present, who is neither Black nor White.
+Thus, $\beta_0 + b_i$, or the city-specific intercept, represents the log odds of being fatally shot for a White, unarmed victim in city $i$ when there is only one officer present, who is neither Black nor White.
 The variabilities of the intercepts and slopes across different cities and different race groups within cities are all greater than 0, indicating that there is enough variation in the baseline odds of being shot fatally in different cities and different race groups within cities.
-From the fixed effect coefficients, we can see that generally, all other race types are less likely to be shot fatally than White 'typical' individuals.
+From the fixed effect coefficients, we can see that generally, all other race types are less likely to be shot fatally than White individuals in 'typical' shootings.
 
@@
 From this table, we can see that the city with the highest baseline odds of being shot fatally is Albuquerque, whereas the city with the lowest baseline odds of being shot fatally is Chicago.
-For White people in Chicago, the predicted log odds of being shot fatally is 0.76 less than that of in a 'typical' scenario and city.
-However, for Black people in Chicago, the predicted log odds of being shot fatally is actually 0.28 higher than that of a 'typical' scenario and city.
-In contrast, for Latino people in Chicago, the predicted log odds of being shot fatally is 0.23 lower than that of a 'typical' scenario and city, and for people of other race types in Chicago, the predicted log odds of being shot fatally is 2.86 lower than that of a 'typical' scenario and city.
-We can conclude that in Chicago, Black people are disproportionately shot fatally in police shootings.
-In Albuquerque, the trend is almost opposite: for Black people, the log odds of being shot fatally is 0.43 less than that of a 'typical' scenario and city, whereas for White people, the log odds of being shot fatally is 0.67 more.
-Overall, we can conclude that city plays a large role in whether a race type is fatally shot in police shootings, although the general trend is that White people are most likely to be shot fatally than other race types.
+For White people in Chicago, the predicted log odds of being shot fatally is 0.76 less than that in a 'typical' shooting.
+However, for Black people in Chicago, the predicted log odds of being shot fatally is actually 0.28 higher than that in a 'typical' shooting.
+In contrast, for Latino people in Chicago, the predicted log odds of being shot fatally is 0.23 lower than that in a 'typical' shooting, and for people of other race types in Chicago, the predicted log odds of being shot fatally is 2.86 lower than that in a 'typical' shooting.
+We can conclude that in Chicago, Black people are disproportionately shot fatally in police shootings with singular victims.
+In Albuquerque, the trend is almost opposite: for Black people, the log odds of being shot fatally is 0.43 less than that in a 'typical' shooting, whereas for White people, the log odds of being shot fatally is 0.67 more.
+Overall, we can conclude that city plays a large role in whether a race type is fatally shot in police shootings, although the general trend is that White individuals are most likely to be shot fatally than other race types.
 
-### Multiple Victims Model
+## Multiple Victims Model
 
 ```{r mult random effects}
-rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
+rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 kable(summary(rand_fitm)$coefficients)
-kable(ranef(rand_fitm))
 ```
 
-|        In this multiple victims model, the $\beta_0$ coefficient, which has a value of -1.77, represents the log odds of being fatally shot by police in a 'typical' scenario and city.
-'Typical' can be defined as a shooting in which the victim is White and unarmed and there is only one officer present, who is neither Black nor Latino.
-Thus, $\beta_0 + b_i$, or the city-specific intercept, represents the log odds of being fatally shot for a 'typical' White, unarmed victim in city $i$ when there is only one officer present, who is neither Black nor White.
+|        In this multiple victims model, the $\beta_0$ coefficient, which has a value of -2.03, represents the log odds of being fatally shot by police in a 'typical' shooting.
+This can be defined as a shooting in which the victim is White and unarmed and there is only one officer present, who is neither Black nor Latino.
+Thus, $\beta_0 + b_i$, or the city-specific intercept, represents the log odds of being fatally shot for a White, unarmed victim in city $i$ when there is only one officer present, who is neither Black nor Latino.
 The variabilities of the intercepts and slopes across different cities and different race groups within cities are all greater than 0, indicating that there is enough variation in the baseline odds of being shot fatally in different cities and different race groups within cities.
-From the fixed effect coefficients, we can see that generally, all other race types are less likely to be shot fatally than White 'typical' individuals.
+From the fixed effect coefficients, we can see that generally, the Black and Other race types are less likely to be shot fatally than White individuals in 'typical' shootings, while Latino individuals are slightly more likely to be shot fatally than these White individuals.
 
-|        The `ranef` function also allows us to quantify the difference between the average predicted log odds of being fatally shot for the general population versus that of a specific city and subject race type, other predictors remaining unchanged.
-This can be found in the Appendix.
-From this table, we can see that the city with the highest baseline odds of being shot fatally is Albuquerque, whereas the city with the lowest baseline odds of being shot fatally is Chicago.
-For White people in Chicago, the predicted log odds of being shot fatally is 0.76 less than that of in a 'typical' scenario and city.
-However, for Black people in Chicago, the predicted log odds of being shot fatally is actually 0.28 higher than that of a 'typical' scenario and city.
-In contrast, for Latino people in Chicago, the predicted log odds of being shot fatally is 0.23 lower than that of a 'typical' scenario and city, and for people of other race types in Chicago, the predicted log odds of being shot fatally is 2.86 lower than that of a 'typical' scenario and city.
-We can conclude that in Chicago, Black people are disproportionately shot fatally in police shootings.
-In Albuquerque, the trend is almost opposite: for Black people, the log odds of being shot fatally is 0.43 less than that of a 'typical' scenario and city, whereas for White people, the log odds of being shot fatally is 0.67 more.
-Overall, we can conclude that city plays a large role in whether a race type is fatally shot in police shootings, although the general trend is that White people are most likely to be shot fatally than other race types.
+|        The output of the `ranef` function for this model can again be found in the Appendix.
+From this table, we can see that the city with the highest baseline odds of being shot fatally is Tucson, whereas the city with the lowest baseline odds of being shot fatally is again Chicago.
+For non-Black and non-Latino people in Chicago, the predicted log odds of being shot fatally is 1.04 less than that in a 'typical' shooting.
+However, for Black people in Chicago, the predicted log odds of being shot fatally is 2.23 higher than that in a 'typical' shooting.
+In contrast, for Latino people in Chicago, the predicted log odds of being shot fatally is 0.70 higher than that in a 'typical' shooting but still lower than that for Black people.
+We can conclude that in Chicago, Black people are disproportionately shot fatally in police shootings with multiple victims.
+In Tucson, the trend is almost opposite: for Black people, the log odds of being shot fatally is 1.37 less than that in a 'typical' shooting, whereas for non-Black and non-Latino people, the log odds of being shot fatally is 0.64 more.
+Overall, we can conclude that city plays a large role in whether a race type is fatally shot in police shootings with multiple victims, although the general trend is that Latino individuals are most likely to be shot fatally than other race types.
 
-### Differences Between Single and Multiple Victims
+## Differences Between Single and Multiple Victim Shootings
 
-### Model Validation and Diagnostics
+|        In both types of shooting, it was extremely important to capture correlations within cities, since there proved to be significant variations in trends between different cities.
+Additionally, it appears that most of the predictors remained the same between the two models, the only difference being that the presence of a Latino officer was statistically significant in multiple-victim shootings, whereas the presence of a White officer was statistically significant in single-victim shootings.
+Unfortunately, I was unable to add a random slope for Other race type individuals in the multiple-victim shootings because it forced the intercept not to be random anymore.
+Thus, I was unable to interpret the multiple-victim shootings coefficients in terms of White vs.
+other race types of individuals.
+
+|        Some trends that remained constant between the types of shootings were that having multiple officers there increased the log odds of the shooting being fatal, whereas having a Black officer present decreased the log odds of the shooting being fatal.
+Additionally, the victim having a knife seemed to lead to the largest log odds of the shooting being fatal out of all types of weapons.
+Finally, many trends within cities remained the same among both types of shootings, one prominent example being Chicago, in which Black individuals are disproportionately shot fatally.
+One trend that changed between the two types of shootings was that White individuals were most likely to be shot fatally than other race types in single-victim shootings, whereas Latino individuals were most likely to be shot fatally in multiple-victim shootings.
+
+## Model Validation and Diagnostics
 
@@
 For the single victims model, I achieved an average 5-fold cross-validation prediction accuracy of 0.66 and a test set prediction accuracy of 0.70.
-For the multiple victims model, I achieved an average 5-fold cross-validation prediction accuracy of 0.61, and a test set prediction accuracy of 0.63.
+For the multiple victims model, I achieved an average 5-fold cross-validation prediction accuracy of 0.63, and a test set prediction accuracy of 0.40.
+This test set prediction accuracy was relatively low, and although adding $\frac{black_i}{white_i}$ into the model significantly increasing the predictive accuracy of the model, I decided not to include it as a predictor because I suspected it was being used as a sort of index to identify the city that the shooting occurred in, which led to such a large boost in test set prediction accuracy.
 
-### Sensitivity Analysis
+|        I will use a binned residual plot and QQ-plot in the next submission for model diagnostics.
 
-## V. Discussion and Conclusion
+## Sensitivity Analysis
 
-|        Overall, 
+|        For sensitivity analysis, I plan to test my assumptions about my imputed values, especially in the `fatal` column.
+Instead of assuming 'U' means 'N', I will instead assume that it means 'Y' and thus will be filtering out less columns as a result.
+I will incorporate this analysis into my next submission.
+
+# V. Discussion and Conclusion
+
+|        Overall, it appears that the most of the statistically significant predictors in single-victim vs.
+multiple-victim shootings overlap, meaning there is not a huge difference in these two types of shootings.
+Furthermore, the general trends among both types of shootings seem to be pretty similar, with the most important trend that changed among the two being that White individuals were most likely to be shot fatally than other race types in single-victim shootings, whereas Latino individuals were most likely to be shot fatally in multiple-victim shootings.
+Finally, it appears that legislative predictors are not significant in predicting whether a shooting will be fatal or not, whereas city, number of officers, officer race, victim weapon-type, and subject race are statistically significant.
+Surprisingly, White people generally have one of the highest odds ratios of being fatally shot in both types of police shootings, although this trend differs by city.
+
 |        One strength of my analysis is that I am considering many variables as predictors, from individual level data to contextual predictors, such as state gun legislation.
@@
 Finally, there were less than 200 cases in which there were multiple victims in my final dataset, which is a relatively small sample size.
-Thus, it may be necessary to take the results from the multiple-victim model with a grain of salt.
+Thus, it may be necessary to take the results from the multiple-victim model with a grain of salt, especially since the test set predictive accuracy was so low.
 
-## VI. References
+# VI. References
 
@@
 
-## VII. Appendix
+# VII. Appendix
 
@@
 
-#### Single-Victim Shooting
+### Single-Victim Shooting
 ```{r single fit}
-rand_fit2 <- glmer(fatal ~ SubjectRace + ob + ow + omult + gun + weapon_other + replica + knife + (SubjectRace | city), data = single_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 kable(ranef(rand_fit2), caption = "Random Effects Coefficients for Single-Victim Shootings")
@@
 
-#### Multiple-Victim Shooting
+### Multiple-Victim Shooting
 ```{r mult fit}
-rand_fitm <- glmer(fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city), data = mult_victims, family = "binomial", control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
 kable(ranef(rand_fitm), caption ="Random Effects Coefficients for Multiple-Victim Shootings")
@@
 
-m0 <- 'fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + I(black/white) + (1 + sb + sl | city)'
+m0 <- 'fatal ~ sb + sl + so + ob + ol + omult + gun + knife + weapon_other + replica + (1 + sb + sl | city)'
 p0 <- crossvalidate(data, k = 5, model = m0, dependent = 'fatal', random = TRUE)
